### 2015-01-07 Project writeup.tex
Adding paper progress

\documentclass[a4paper,12pt]{article}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\usepackage{amsthm}

\newtheorem{claim}{Claim}

\begin{document}

\title{Estimating Trending Topics on Twitter with Small Subsets of the Total Data}

\author{Albert Lee, Evan Miller, and Kiran Vodrahalli}

\maketitle

\section{Introduction}

\section{Problem}

\section{Previous Work}

\section{Algorithm}

\subsection{Description}

We separate the problem of finding trending topics on Twitter into two parts.  First, we need to maintain a data structure that efficiently stores data about all occurrences of every hashtag seen in the past.  We also maintain a separate data structure that allows us to quickly gather information about the most recent hashtags seen.

We want the former data structure to be very space efficient since it must store data about a very large dataset.  For this structure, space efficiency is more important than accuracy since small deviations in such a large dataset should not be significant because these deviations in past data should not greatly effect what is currently trending.

For the latter data structure, accuracy is more important than space efficiency since the structure contains data which more closely relates to which topics are currently trending and the size of the dataset is much smaller.

\subsubsection{Data structure for past tweets}

To store data about all occurrences of every hashtag seen in the past, we use a modified version of the time-aggregated Hokusai system\footnote{http://arxiv.org/pdf/1210.4891v1.pdf}, which is an extension of the Count-Min sketch.  This data structure is described in full detail in section \ref{Hokusai}.

The goal of the time-aggregated Hokusai system is to store older data at decreased precision since the older data also has decreased value in the computation.  The time-aggregated Hokusai system works by storing aggregate data in Count-Min sketches with a $2^i$ day resolution for the past $2^i$ days.  Each of these Count-Min sketches computes $m$ hash functions from $\{0,1\}^*$ to $\{0, 1, …, n-1\}$.

To this structure we add another layer that combines these Count-Min sketches into a single Count-Min sketch that depreciates the value of older data.  This aggregate Count-Min sketch stores the values $\bar{M} + \sum\limits_{j=1}^{\log⁡(T)} \frac{M^j}{2^j}$ where $\bar{M}$ is the Count-Min sketch storing today's data, $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ resolution, and $T$ is an upper bound on the number of days stored by the data structure.  In section \ref{Correct} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i$ days ago with value approximately $\frac{1}{2i}$.

\subsubsection{Data structure for current tweets}

Our data structure for holding hashtags seen in the last $y$ minutes consists of three components: a max Fibonacci heap, a queue, and a hash table.

The keys for the hash table are the hashtags, and the values stored in the table are (frequency, pointer to corresponding node in heap) pairs.

The queue contains (hashtag, timestamp) pairs, each of which is inserted upon seeing a hashtag in the input stream.

The heap has keys equal to (frequency in last $y$ minutes) / (value in Hokusai data structure) for a specific hashtag, and the value stored in the node is the corresponding hashtag.

\subsubsection{Updating Hokusai data structure\label{Hokusai}}

Algorithm \ref{alg1} describes the necessary steps to maintain the Hokusai data structure as new input is provided.

\begin{algorithm}
\caption{Time Aggregation for the Hokusai structure}\label{alg1}
\begin{algorithmic}[1]
\ForAll {$i$}
	\State Initialize Count-Min sketch $M^i = 0$
\EndFor
\State Initialize $t = 0$, $\bar{M} = 0$, and $A = 0$
\While {data arrives}
	\State Aggregate data into sketch $\bar{M}$ for current day while also adding this data to $A$
	\State $t \leftarrow t + 1$ (increment counter)
	\State $A \leftarrow A - \bar{M}$
	\For {$j = 0$ to argmax \{$l$ where $t$ mod $2^l = 0$\}}
		\State $A \leftarrow A + 2^{-j}(\bar{M} - M^j)$
		\State $T \leftarrow \bar{M}$ (back up temporary storage)
		\State $\bar{M} \leftarrow \bar{M} + M^j$ (increment cumulative sum)
		\State $M^j \leftarrow T$ (new value for $M^j$)
	\EndFor
	\State $\bar{M} \leftarrow 0$ (reset aggregator)
\EndWhile
\end{algorithmic}
\end{algorithm}

\subsubsection{Updating heap and hash tables}

\begin{algorithm}
\caption{Thread to process data input.}
\begin{algorithmic}[1]
\While {data arrives}
	\If {hashtag is in hash table}
		\State Increment the frequency stored at that entry.
		\State Update the key of the node in the Fibonacci heap.
	\Else
		\State Insert a new node into the Fibonacci heap with the appropriate key and value. 
		\State Insert the hashtag into the hash table with a pointer to this node.
 	\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Thread to discard oldest recent hashtags.}
\begin{algorithmic}[1]
\Loop
	\If {queue is not empty}
		\State Peek at end of queue.
		\If {the timestamp + $y$ minutes is before the current time}
			\State Look up the hashtag in the hash table and decrement the stored frequency.
			\If {the frequency is now 0}
				\State Delete the node in the heap pointed to by this entry in the table.
				\State Delete this entry in the hash table.
			\Else	
				\State Update the key of this node pointed to by this entry in the table to the proper value given the new frequency.
			\EndIf
		\EndIf
	\EndIf
\EndLoop
\State Make the hash table for the oldest $x$ minute interval correspond to the upcoming $x$ minute interval.

\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Updates at end of a day.}
\begin{algorithmic}[1]
\State Do all the end-of-day aggregation for the Hokusai structure as detailed in Algorithm \ref{alg1}.
\ForAll {elements in the Fibonacci heap}
	\State look up the hashtag corresponding to this node in the hash table
	\State Update the key of the node to: \[\frac{\text{frequency in last y minutes found at the table entry}}{\text{new value in Hokusai data structure}}\]
\EndFor
\end{algorithmic}
\end{algorithm}

\subsubsection{Finding the trending hashtags}

\begin{algorithm}
\caption{Algorithm to find top $k$ trending items.}
\begin{algorithmic}[1]
\State Perform $k$ delete-max operations on the Fibonacci heap, storing each of the nodes deleted in a list $L$.
\ForAll {nodes $N$ in $L$}
	\State Announce that the hashtag associated with $N$ is trending.
	\State Insert $N$ into the Fibonacci heap.
\EndFor
\end{algorithmic}
\end{algorithm}

\subsection{Analysis}

Maybe put table here of all tunable parameters mentioned in the paper to remind readers what they refer to.

\subsubsection{Time analysis}

Let s = size of heap = number of distinct hashtags seen in last y minutes.

Time for actions when a hashtag comes in:

	Takes amortized O(m) time to update Hokusai structure.

	Takes expected O(1) time to check if in hash table.  If so:

		O(1) time to increment

		O(m) time to compute key

		O(1) amortized time to update key since key nondecreasing

	Else:

		O(m) time to compute key value and O(1) time to insert node in heap

		Takes O(1) time to insert into hash table.

Thus, takes O(m) amortized time + expected O(1) time to do the actions when a hashtag comes in.

Time for actions when $y$ minutes pass after a hashtag comes in:
	O(1) time to check if queue is empty.  If not:
		O(1) time to look at the queue and see if the timestamp + $y$ minutes is before the current time.  If so:
			Expected O(1) time to look up in hash table and decrement the frequency.
			O(1) time to check if frequency is 0.  If so:
				O(log(s)) amortized time to delete node in heap.
				O(1) time to delete entry in hash table.
			Else:
				O(m) amortized time to compute the 
				O(log(s)) amortized time to update the heap given this key.

Thus, takes O(log(s)) amortized time + expected O(1) time  + O(m) time to do actions when a hashtag is discarded.

Time for actions when day ends:
Takes O(nm) amortized time to do all end of day aggregation.
Takes O(m) time to compute each key.
Takes O(1) amortized time to update each key since all keys are nondecreasing.

Thus, takes O(nm + s) amortized time + O(ms) time at the end of the day.

Time for actions when finding top k trending items:
	Takes O(k*log(s)) time for delete-max operations.
	Takes O(k) time to announce these as trending.
Takes O(k) amortized time to reinsert these nodes into the heap.

Thus, takes O(k*log(s)) time + O(k) amortized time to determine what’s trending.

\subsubsection{Space analysis}

Hokusai structure takes O(nm*log(T)) space.

Heap takes O(s) space.

Hash table needs O(s) space to keep expected O(1) lookup time.

Queue requires O(t) space. (where t max hashtags seen in y min interval)

Thus, everything requires O(nm*log(T) + s + t) space.

\subsubsection{Correctness \label{Correct}}

This algorithm finds the k hashtags that have the maximum value of (frequency in last y minutes) / (value in Hokusai data structure).

\begin{claim}
The value for hashtag $x$ in the aggregate Count-Min sketch of the Hokusai data structure is within a factor of 4 of $\bar{M}(x) + \sum\limits_{i=1}^T \frac{1}{i}*$ (value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring i days ago).
\end{claim}
\begin{proof}
First, we use Theorem 4 of the Hokusai paper\footnote{http://arxiv.org/pdf/1210.4891v1.pdf} which states that ``At $t$, the sketch $M^j$ contains statistics for the period $[t - \delta, t - \delta - 2^j]$ where $\delta = t \mod 2^j$.''

Let $b$ be the location in the aggregate Count-Min sketch containing the value returned when $x$ is queried.

Let $h$ be any instance of any hashtag that appeared on day $m$ such that seeing $h$ incremented counters in position $b$.

Case 1: $m=t$

Then the statistics for $h$ are recorded in $\bar{M}$ and are not in any $M^j$.

Case 2: $2^i > t-m \geq 2^{i-1}$ for some $i > 0$

By Theorem 4, for all $j \leq i - 2$, $M^j$ does not contain statistics for $m$ since $m \leq t - 2^{i-1} \leq t - \delta - 2^{i-2}$.

Therefore, the increments that occurred in the Count-Min sketch for hashtags occurring $i$ days ago contribute at most $\sum\limits_{j = i - 1}^{T} 2^{-j} < 2^{2-i}$ to the value in position $b$.

Let $k$ be the largest $j$ such that $t - \delta - 2^j \geq m$

Then $m \leq t - \delta - 2^k$.  Let $\lambda = t \mod 2^{k+1}$.  Then $\lambda = \delta$ or $\lambda = \delta + 2^k$.

Since $k$ is the largest $j$ such that $t - \delta - 2^j \geq m$, $m > t - \lambda - 2^{k+1}$.

Also, $m \leq t - \delta - 2^k \leq t - \lambda$, so $M^{k+1}$ contains statistics about $h$.

For all $j \geq i$, $t - \delta - 2^j < t - 2^j < m$, so $k \leq i - 1$.

Thus, incrementing the counter in $M^{k + 1}$ contributed at least $2^{-i}$ to the value in position $b$.

Thus, the contributions to the sum are within a factor of 4 of $\frac{1}{t-m}$.

Therefore, summing over all hashtags that increment counters in position $b$ gives $\bar{M}(x)$ for all hashtags that occurred on day $t$, and within a factor of 4 of $\sum\limits_{i=1}^T \frac{1}{i}*$(value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring $i > 0$ days ago). 
\end{proof}
This value is approximately (frequency in last y minutes) / (freq today + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag i days ago)).
This seems to be a desirable function to maximize since it finds hashtags that are common in the last y minutes that have been comparatively infrequent in the past.  This function is good since it especially emphasizes hashtags that are different than those seen in the past few days.  This ensures that the same items do not stay trending for too long.

\section{Experiments}

\section{Variants}

\subsection{Heap function}

\subsection{Modify heap to always know what's trending}

(Good for accuracy, bad since forfeits gain from using Fibonacci heap since items will be deleted from tree more frequently.)

\subsection{Deamortizing end of day cost}
\end{document}

### 2015-01-07 fib_heap.py
added paper + fib_heap + tweet_trender (non-naive implementation)

@@ -1,0 +1,638 @@
+
+'''
+File: fibonacci_heap_mod.py
+Author: Keith Schwarz (htiek@cs.stanford.edu)
+Ported to Python by Dan Stromberg (strombrg@gmail.com)
+
+An implementation of a priority queue backed by a Fibonacci heap, as described
+by Fredman and Tarjan.  Fibonacci heaps are interesting theoretically because
+they have asymptotically good runtime guarantees for many operations.  In
+particular, insert, peek, and decrease-key all run in amortized O(1) time.
+dequeue_min and delete each run in amortized O(lg n) time.  This allows
+algorithms that rely heavily on decrease-key to gain significant performance
+boosts.  For example, Dijkstra's algorithm for single-source shortest paths can
+be shown to run in O(m + n lg n) using a Fibonacci heap, compared to O(m lg n)
+using a standard binary or binomial heap.
+
+Internally, a Fibonacci heap is represented as a circular, doubly-linked list
+of trees obeying the min-heap property.  Each node stores pointers to its
+parent (if any) and some arbitrary child.  Additionally, every node stores its
+degree (the number of children it has) and whether it is a "marked" node.
+Finally, each Fibonacci heap stores a pointer to the tree with the minimum
+value.
+
+To insert a node into a Fibonacci heap, a singleton tree is created and merged
+into the rest of the trees.  The merge operation works by simply splicing
+together the doubly-linked lists of the two trees, then updating the min
+pointer to be the smaller of the minima of the two heaps.  Peeking at the
+smallest element can therefore be accomplished by just looking at the min
+element.  All of these operations complete in O(1) time.
+
+The tricky operations are dequeue_min and decrease_key.  dequeue_min works by
+removing the root of the tree containing the smallest element, then merging its
+children with the topmost roots.  Then, the roots are scanned and merged so
+that there is only one tree of each degree in the root list.  This works by
+maintaining a dynamic array of trees, each initially null, pointing to the
+roots of trees of each dimension.  The list is then scanned and this array is
+populated.  Whenever a conflict is discovered, the appropriate trees are merged
+together until no more conflicts exist.  The resulting trees are then put into
+the root list.  A clever analysis using the potential method can be used to
+show that the amortized cost of this operation is O(lg n), see "Introduction to
+Algorithms, Second Edition" by Cormen, Rivest, Leiserson, and Stein for more
+details.
+
+The other hard operation is decrease_key, which works as follows.  First, we
[…]

### 2015-01-07 tweet_trender.py

@@ -1,0 +1,23 @@
+# COS 521 Final Project
+# Algorithm Implementation 
+
+import fib_heap as fh
+
+
+# current tweet storage
+
+# queue for holding hashtags from the stream
+# (hashtag, timestamp pairs)
+input_queue = []
+
+# hash table
+# map from hashtag to (frequency, ptr to heap) pairs
+hashtag_freq = dict()
+
+# max Fib heap
+# decrease_key is amortized constant
+# To benefit from Fibonacci heaps in practice, 
+# you have to use them in an application where decrease_keys are incredibly frequent.
+# we want to do k delete-max ops in the heap 
+# http://stromberg.dnsalias.org/~strombrg/fibonacci-heap-mod/
+print "this worked again"

### 2015-01-07 key_functions.py
few updates

@@ -1,0 +1,5 @@
+# COS 521 Final Project
+# key functions used for Hokusai
+
+
+

### 2015-01-07 naive_trend_predictor.py

@@ -10,7 +10,7 @@ class NaiveTrendPredictor:
         with open(data_file) as f:
             for line in f:
                 try:
-                    id, timestamp, hashtag = line.strip().split(',')
+                    id_, timestamp, hashtag = line.strip().split(',')
                     tweet_dt = parser.parse(timestamp)
                 except Exception, e:
                     print e

### 2015-01-07 tweet_trender.py

@@ -2,22 +2,58 @@
 # Algorithm Implementation 
 
 import fib_heap as fh
+from collections import default_dict
+
+import key_functions
 
 
-# current tweet storage
+class SmartTrendPredictor:
+	# update the hashtag frequency dict and heap data structures
+	# new_data is in the format (hashtag, timestamp)
+	def update_datastructs(new_data):
+		hashtag, timestamp = new_data
+		if hashtag in self.hashtag_freq:
+			# update the frequency
+			self.hashtag_freq[hashtag][0] += 1
+		else:
+			# build the ptr to the heap
+			# calculate key
+			self.heap.enqueue()
 
-# queue for holding hashtags from the stream
-# (hashtag, timestamp pairs)
-input_queue = []
+	def __init__(self, data_file):
+		self.tweets = defaultdict(list)
+		self.inverted_tweets_idx = defaultdict(list)
+		# current tweet storage
+		'''
+		# queue for holding hashtags from the stream
+		# (hashtag, timestamp pairs)
+		self.input_queue = []
+		'''
+		# map from hashtag to (frequency, ptr to heap) pairs
+		self.hashtag_freq = dict()
+		# max Fib heap
+		# decrease_key is amortized constant
+		# To benefit from Fibonacci heaps in practice, 
+		# you have to use them in an application where decrease_keys are incredibly frequent.
+		# we want to do k delete-max ops in the heap 
+		# http://stromberg.dnsalias.org/~strombrg/fibonacci-heap-mod/
+		# this is a min heap, so enter all keys as negative to get the max heap.
+		self.heap = fh.Fibonacci_heap()
 
-# hash table
-# map from hashtag to (frequency, ptr to heap) pairs
-hashtag_freq = dict()
 
-# max Fib heap
-# decrease_key is amortized constant
-# To benefit from Fibonacci heaps in practice, 
-# you have to use them in an application where decrease_keys are incredibly frequent.
-# we want to do k delete-max ops in the heap 
-# http://stromberg.dnsalias.org/~strombrg/fibonacci-heap-mod/
-print "this worked again"
+		# Hokusai data structure 
+		self.hokusai = 0
+
+		with open(data_file) as f:
+			for line in f:
+				try:
+	                id_, timestamp, hashtag = line.strip().split(',')
+	                tweet_dt = parser.parse(timestamp)
+	            except Exception, e:
+	                print e
+	                continue
[…]

### 2015-01-07 tweet_archive_parser.py
Add parser for twitter archive files

@@ -1,0 +1,29 @@
+import sys
+import json
+import codecs
+
+def parse_tweet_file(tweet_file):
+    '''Writes a file of
+            tweet_id,timestamp,hashtag
+       for the tweets in the input json file.'''
+    outfile = tweet_file.replace('.json', '.out')
+    out_f = codecs.open(outfile, 'w', encoding='utf-8')
+    with open(tweet_file) as f:
+        for line in f:
+            tweet = json.loads(line.strip())
+            if ('id' in tweet and 'created_at' in tweet and
+                'text' in tweet and 'lang' in tweet and
+                tweet['lang'] == 'en'):
+                if '#' in tweet['text']:
+                    tweet_id = unicode(tweet['id'])
+                    tweet_timestamp = tweet['created_at']
+                    tag_tokens = tweet['text'].strip().split('#')[1:]
+                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens]
+                    for tag in tags:
+                        out_f.write(','.join([tweet_id, tweet_timestamp, tag+'\n']))
+
+if __name__=='__main__':
+    if len(sys.argv) != 2:
+        print 'Usage: python tweet_parser.py <tweet_file.json>'
+        sys.exit()
+    parse_tweet_file(sys.argv[1])

### 2015-01-07 tweet_parser.py
Parser for dirs of tweet archives

@@ -1,0 +1,36 @@
+import sys
+import json
+import codecs
+
+def parse_tweet_file(tweet_file):
+    '''Writes a file of
+            tweet_id,timestamp,hashtag
+       for the tweets in the input json file.'''
+    outfile = tweet_file.replace('.json', '.out')
+    out_f = codecs.open(outfile, 'w', encoding='utf-8')
+    with open(tweet_file) as f:
+        for line in f:
+            tweet = json.loads(line.strip())
+            if ('id' in tweet and 'created_at' in tweet and
+                'text' in tweet and 'lang' in tweet and
+                tweet['lang'] == 'en'):
+                if '#' in tweet['text']:
+                    tweet_id = unicode(tweet['id'])
+                    tweet_timestamp = tweet['created_at']
+                    tag_tokens = tweet['text'].strip().split('#')[1:]
+                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens]
+                    for tag in tags:
+                        out_f.write(','.join([tweet_id, tweet_timestamp, tag+'\n']))
+
+
+if __name__=='__main__':
+    if len(sys.argv) != 2:
+        print 'Usage: python tweet_parser.py <DATA_DIR>'
+        sys.exit()
+
+    import os
+    for root, dirs, files in os.walk(sys.argv[1]):
+        for filename in files:
+            if filename.endswith('.json'):
+                print os.path.join(root, filename)
+                parse_tweet_file(os.path.join(root, filename))

### 2015-01-08 tweet_parser.py
Fix parser for double hashtags

@@ -18,7 +18,7 @@ def parse_tweet_file(tweet_file):
                     tweet_id = unicode(tweet['id'])
                     tweet_timestamp = tweet['created_at']
                     tag_tokens = tweet['text'].strip().split('#')[1:]
-                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens]
+                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens if token]
                     for tag in tags:
                         out_f.write(','.join([tweet_id, tweet_timestamp, tag+'\n']))
 

### 2015-01-08 tweet_parser.py
Fix tweet parser

@@ -18,7 +18,7 @@ def parse_tweet_file(tweet_file):
                     tweet_id = unicode(tweet['id'])
                     tweet_timestamp = tweet['created_at']
                     tag_tokens = tweet['text'].strip().split('#')[1:]
-                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens if token]
+                    tags = [token.split()[0].rstrip('?:!.,;') for token in tag_tokens if token.split()]
                     for tag in tags:
                         out_f.write(','.join([tweet_id, tweet_timestamp, tag+'\n']))
 

### 2015-01-10 countminsketch.py
added count min sketch

@@ -1,0 +1,94 @@
+## DOWNLOADED FROM 
+## https://github.com/rafacarrascosa/countminsketch
+## (available on pip)
+
+# -*- coding: utf-8 -*-
+import hashlib
+import array
+
+
+class CountMinSketch(object):
+    """
+    A class for counting hashable items using the Count-min Sketch strategy.
+    It fulfills a similar purpose than `itertools.Counter`.
+
+    The Count-min Sketch is a randomized data structure that uses a constant
+    amount of memory and has constant insertion and lookup times at the cost
+    of an arbitrarily small overestimation of the counts.
+
+    It has two parameters:
+     - `m` the size of the hash tables, larger implies smaller overestimation
+     - `d` the amount of hash tables, larger implies lower probability of
+           overestimation.
+
+    An example usage:
+
+        from countminsketch import CountMinSketch
+        sketch = CountMinSketch(1000, 10)  # m=1000, d=10
+        sketch.update("oh yeah")
+        sketch.update(tuple())
+        sketch.update(1, value=123)
+        print sketch["oh yeah"]       # prints 1
+        print sketch[tuple()]         # prints 1
+        print sketch[1]               # prints 123
+        print sketch["non-existent"]  # prints 0
+
+    Note that this class can be used to count *any* hashable type, so it's
+    possible to "count apples" and then "ask for oranges". Validation is up to
+    the user.
+    """
+
+    def __init__(self, m, d):
+        """ `m` is the size of the hash tables, larger implies smaller
+        overestimation. `d` the amount of hash tables, larger implies lower
+        probability of overestimation.
+        """
+        if not m or not d:
+            raise ValueError("Table size (m) and amount of hash functions (d)"
+                             " must be non-zero")
+        self.m = m
+        self.d = d
+        self.n = 0
+        self.tables = []
+        for _ in xrange(d):
+            table = array.array("l", (0 for _ in xrange(m)))
+            self.tables.append(table)
+
+    def _hash(self, x):
+        md5 = hashlib.md5(str(hash(x)))
+        for i in xrange(self.d):
+            md5.update(str(i))
+            yield int(md5.hexdigest(), 16) % self.m
+
+    def add(self, x, value=1):
+        """
+        Count element `x` as if had appeared `value` times.
+        By default `value=1` so:
+
+            sketch.add(x)
+
+        Effectively counts `x` as occurring once.
+        """
+        self.n += value
+        for table, i in zip(self.tables, self._hash(x)):
+            table[i] += value
+
+    def query(self, x):
[…]

### 2015-01-10 tweet_trender.py

@@ -2,9 +2,19 @@
 # Algorithm Implementation 
 
 import fib_heap as fh
+from countminsketch as CountMinSketch
 from collections import default_dict
 
 import key_functions
+
+
+
+
+class Hokusai:
+
+	def __init__(blah):
+		return 'lol'
+
 
 
 class SmartTrendPredictor:

### 2015-01-10 countminsketch.py
updated hokusai; more comments

@@ -7,6 +7,23 @@ import hashlib
 import array
 
 
+# to note: take the min: we get a lower bound; taking max we would get upper bound
+
+# what happens to the case when we have something we've never seen before, but it
+# hashes? how is performance affected here?
+
+# modify to take into account the ter "b" -- as in M_b
+# for different locations in time, or whatever?
+
+# we can compress time intervals together into one count min sketch (with less accuracy)
+# as we go further into the past, so that we don't need to maintain all of it
+# (we still retain approximate counts, we just push them together)
+
+# resolution means "the amount of time for which we are counting the data"
+# 2^m minute resolution means that we're keeping approximate frequency counts
+# for 2^m minutes 
+
+
 class CountMinSketch(object):
     """
     A class for counting hashable items using the Count-min Sketch strategy.

### 2015-01-10 hokusai.py

@@ -1,0 +1,33 @@
+# COS 521 Final Project
+# Hokusai Implementation
+
+# produced from http://www.auai.org/uai2012/papers/231.pdf
+
+# note that Corollary 2 is what allows us to add CM-sketches
+# from different periods of time (that are disjoint)
+
+# time interval lengths of powers of two match up since they
+# are disjoint from each other
+
+
+# we can compress time intervals together into one count min sketch (with less accuracy)
+# as we go further into the past, so that we don't need to maintain all of it
+# (we still retain approximate counts, we just push them together)
+
+# resolution means "the amount of time for which we are counting the data"
+# 2^m minute resolution means that we're keeping approximate frequency counts
+# for 2^m minutes 
+
+# building Hokusai means that we keep a number of countmin sketches with variable 
+# resolution at the start: (1, 2, 4, 8, 16, ..., 2^n) where n is the number of countmin sketches
+# we maintain in the hokusai datastructure. 
+# as time gets further along, the resolutions increase in size. (i.e. 2^n -> larger and so on)
+# we justify this because older frequency data is less important (exponentially) (section 3.1 description)
+# to the current time interval we're trying to get the frequency info of
+# (even more relevant to the trending discussion) 
+
+# WE RETAIN 2^m minute resolution for the past 2^m minutes
+
+
+class Hokusai(object):
+

### 2015-01-10 tweet_trender.py

@@ -10,13 +10,6 @@ import key_functions
 
 
 
-class Hokusai:
-
-	def __init__(blah):
-		return 'lol'
-
-
-
 class SmartTrendPredictor:
 	# update the hashtag frequency dict and heap data structures
 	# new_data is in the format (hashtag, timestamp)

### 2015-01-10 naive_trend_predictor.py
Naive bug fix

@@ -22,7 +22,7 @@ class NaiveTrendPredictor:
         hashtag_grps = (self.inverted_tweets_idx[tweet_dt]
                         for tweet_dt in self.inverted_tweets_idx
                         if start_dt <= tweet_dt and tweet_dt <= end_dt)
-        most_popular = {(len(self.tweets[hashtag]), hashtag)
+        most_popular = {(sum((start_dt <= tweet_dt and tweet_dt <= end_dt) for tweet_dt in self.tweets[hashtag]), hashtag)
                         for hashtag_grp in hashtag_grps
                         for hashtag in hashtag_grp}
         return list(itertools.islice(reversed(sorted(most_popular)), 0, n))

### 2015-01-11 countminsketch.py
updated hokusai part more

@@ -1,6 +1,7 @@
 ## DOWNLOADED FROM 
 ## https://github.com/rafacarrascosa/countminsketch
 ## (available on pip)
+## MODIFIED BY KIRAN VODRAHALLI for purposes of hokusai
 
 # -*- coding: utf-8 -*-
 import hashlib
@@ -15,13 +16,6 @@ import array
 # modify to take into account the ter "b" -- as in M_b
 # for different locations in time, or whatever?
 
-# we can compress time intervals together into one count min sketch (with less accuracy)
-# as we go further into the past, so that we don't need to maintain all of it
-# (we still retain approximate counts, we just push them together)
-
-# resolution means "the amount of time for which we are counting the data"
-# 2^m minute resolution means that we're keeping approximate frequency counts
-# for 2^m minutes 
 
 
 class CountMinSketch(object):
@@ -54,6 +48,9 @@ class CountMinSketch(object):
     possible to "count apples" and then "ask for oranges". Validation is up to
     the user.
     """
+    # make m, d accessible
+    m = 0
+    d = 0
 
     def __init__(self, m, d):
         """ `m` is the size of the hash tables, larger implies smaller
@@ -70,6 +67,17 @@ class CountMinSketch(object):
         for _ in xrange(d):
             table = array.array("l", (0 for _ in xrange(m)))
             self.tables.append(table)
+
+
+    # expose the internal array to update for purposes of addition
+    # - Kiran 
+    # update table i, index j
+    # i in [0, m); j in [0, d)
+    def update(self, i, j, new_val):
+        self.tables[i][j] = new_val
+    # get val at i j
+    def val_at(self, i, j):
+        return self.tables[i][j]
 
     def _hash(self, x):
         md5 = hashlib.md5(str(hash(x)))

### 2015-01-11 hokusai.py

@@ -19,15 +19,74 @@
 # for 2^m minutes 
 
 # building Hokusai means that we keep a number of countmin sketches with variable 
-# resolution at the start: (1, 2, 4, 8, 16, ..., 2^n) where n is the number of countmin sketches
+# resolution: (1, 2, 4, 8, 16, ..., 2^m) where m is the number of countmin sketches
 # we maintain in the hokusai datastructure. 
-# as time gets further along, the resolutions increase in size. (i.e. 2^n -> larger and so on)
+# WE RETAIN 2^m minute resolution for the past 2^m minutes
 # we justify this because older frequency data is less important (exponentially) (section 3.1 description)
 # to the current time interval we're trying to get the frequency info of
 # (even more relevant to the trending discussion) 
 
-# WE RETAIN 2^m minute resolution for the past 2^m minutes
+from countminsketch as CountMinSketch
+from copy import deepcopy
 
+
+# sum two disjoint count-min sketches
+# put the sum into M1, leave M2 alone
+def sketch_sum(M1, M2):
+	if M1.m != M2.m:
+		print "Sketches don't align.\n"
+	elif M1.d != M2.d:
+		print "Sketches don't align.\n"
+	else:
+		for i in range(M2.m):
+			for j in range(M2.d):
+				new_val = M2.val_at(i, j)
+				M1.update(i, j, new_val)
 
 class Hokusai(object):
+	def __init__(self, n, m, d):
+		# number of the CM sketches filled so far
+		# <= n
+		self.num_filled = 0
+		# n is number of CM sketches
+		self.n_ = n 
+		# m is number of hash functions
+		self.m_ = m
+		# d is size of array for each hash function
+		self.d_ = d
+		# n count-min sketches 
+		# we retain resolutions 1, 2, 4, ..., 2^n
+		# contains pairs of (time unit filled, sketch)
+		# move to next sketch (update curr_sketch) when 
+		# time unit filled = 2^i (its position in the list)
+		self.cm_sketch_train = []
+		for i in range(n):
+			self.cm_sketch_train.append([0, CountMinSketch(m, d)])
 
+	# data_block is a block of data
+	# gathered over a time unit
+	def aggregate_unit(self, data_block):
+		# we use this to keep track of the current time unit
+		M_bar = CountMinSketch(m, d)
+		# we use self.M_bar
+		if self.num_filled < self.n_:
+			self.num_filled += 1
[…]

### 2015-01-11 tweet_trender.py

@@ -2,11 +2,17 @@
 # Algorithm Implementation 
 
 import fib_heap as fh
-from countminsketch as CountMinSketch
 from collections import default_dict
 
 import key_functions
 
+# when parsing tweets, need to make sure 
+# the tweet sending model (from file)
+# takes into account times. 
+
+# (so we fake the time sending, basically)
+# (if we want to actually implement realtime,
+#  need a time counter while getting tweets)
 
 
 

### 2015-01-11 countminsketch.py
fixed and mostly finished hokusai stuff; renamed to history"

@@ -65,7 +65,10 @@ class CountMinSketch(object):
         self.n = 0
         self.tables = []
         for _ in xrange(d):
-            table = array.array("l", (0 for _ in xrange(m)))
+            # KIRAN'S EDIT: change "l" to "f" -- we want to allow
+            # float values since we use it to also maintain
+            # weighted counts (and still want to use the min procedure)
+            table = array.array("f", (0 for _ in xrange(m)))
             self.tables.append(table)
 
 

### 2015-01-11 history.py

@@ -1,0 +1,194 @@
+# COS 521 Final Project
+# Implementation of the history data structure
+# Author: Kiran Vodrahalli
+# using Hokusai paper as inspiration with a few modifications
+# citation: http://www.auai.org/uai2012/papers/231.pdf
+
+# note that Corollary 2 is what allows us to add CM-sketches
+# from different periods of time (that are disjoint)
+
+# time interval lengths of powers of two match up since they
+# are disjoint from each other
+
+
+# we can compress time intervals together into one count min sketch (with less accuracy)
+# as we go further into the past, so that we don't need to maintain all of it
+# (we still retain approximate counts, we just push them together)
+
+# resolution means "the amount of time for which we are counting the data"
+# 2^m minute resolution means that we're keeping approximate frequency counts
+# for 2^m minutes 
+
+# building Hokusai means that we keep a number of countmin sketches with variable 
+# resolution: (1, 2, 4, 8, 16, ..., 2^m) where m is the number of countmin sketches
+# we maintain in the hokusai datastructure. 
+# WE RETAIN 2^m minute resolution for the past 2^m minutes
+# we justify this because older frequency data is less important (exponentially) (section 3.1 description)
+# to the current time interval we're trying to get the frequency info of
+# (even more relevant to the trending discussion) 
+
+# we add an extra CMSketch on top of Hokusai to keep track of our weighted sums
+
+from countminsketch import CountMinSketch as CMSketch
+from copy import deepcopy
+
+
+# sum two disjoint count-min sketches
+# put the sum into M1, leave M2 alone
+def sketch_sum(M1, M2):
+	if M1.m != M2.m:
+		print "Sketches don't align on hashtable length.\n"
+		return
+	elif M1.d != M2.d:
+		print "Sketches don't align on # of hashtables.\n"
+		return
+	else:
+		result_sketch = CMSketch(M2.m, M2.d)
+		for i in range(M2.d):
+			for j in range(M2.m):
+				new_val = M1.val_at(i, j) + M2.val_at(i, j)
+				result_sketch.update(i, j, new_val)
+		return result_sketch
+
+
+# multiply each value in the sketch by some factor
+# we don't update the sketch, we return a new one
[…]

### 2015-01-11 hokusai.py

@@ -1,92 +1,0 @@
-# COS 521 Final Project
-# Hokusai Implementation
-
-# produced from http://www.auai.org/uai2012/papers/231.pdf
-
-# note that Corollary 2 is what allows us to add CM-sketches
-# from different periods of time (that are disjoint)
-
-# time interval lengths of powers of two match up since they
-# are disjoint from each other
-
-
-# we can compress time intervals together into one count min sketch (with less accuracy)
-# as we go further into the past, so that we don't need to maintain all of it
-# (we still retain approximate counts, we just push them together)
-
-# resolution means "the amount of time for which we are counting the data"
-# 2^m minute resolution means that we're keeping approximate frequency counts
-# for 2^m minutes 
-
-# building Hokusai means that we keep a number of countmin sketches with variable 
-# resolution: (1, 2, 4, 8, 16, ..., 2^m) where m is the number of countmin sketches
-# we maintain in the hokusai datastructure. 
-# WE RETAIN 2^m minute resolution for the past 2^m minutes
-# we justify this because older frequency data is less important (exponentially) (section 3.1 description)
-# to the current time interval we're trying to get the frequency info of
-# (even more relevant to the trending discussion) 
-
-from countminsketch as CountMinSketch
-from copy import deepcopy
-
-
-# sum two disjoint count-min sketches
-# put the sum into M1, leave M2 alone
-def sketch_sum(M1, M2):
-	if M1.m != M2.m:
-		print "Sketches don't align.\n"
-	elif M1.d != M2.d:
-		print "Sketches don't align.\n"
-	else:
-		for i in range(M2.m):
-			for j in range(M2.d):
-				new_val = M2.val_at(i, j)
-				M1.update(i, j, new_val)
-
-class Hokusai(object):
-	def __init__(self, n, m, d):
-		# number of the CM sketches filled so far
-		# <= n
-		self.num_filled = 0
-		# n is number of CM sketches
-		self.n_ = n 
-		# m is number of hash functions
-		self.m_ = m
-		# d is size of array for each hash function
-		self.d_ = d
-		# n count-min sketches 
-		# we retain resolutions 1, 2, 4, ..., 2^n
-		# contains pairs of (time unit filled, sketch)
-		# move to next sketch (update curr_sketch) when 
-		# time unit filled = 2^i (its position in the list)
-		self.cm_sketch_train = []
-		for i in range(n):
[…]

### 2015-01-11 tweet_trender.py
updated main a bit

@@ -3,7 +3,7 @@
 
 import fib_heap as fh
 from collections import default_dict
-
+import history as h
 import key_functions
 
 # when parsing tweets, need to make sure 

### 2015-01-11 history.py
updated main smartpredictor file

@@ -151,6 +151,9 @@ class History(object):
 		# reset the present now that we're done with one time block
 		self.present = CMSketch(self.m, self.d)
 
+
+	# CURRENTLY THERE IS A MISMATCH BETWEEN THE TWO-- BUG
+	
 	# we want to put these values into its own count-min sketch, (call it A)
 	# updated in sync so as to not waste log T time summing
 	# for each query.

### 2015-01-11 tweet_trender.py

@@ -15,12 +15,40 @@ import key_functions
 #  need a time counter while getting tweets)
 
 
-
+# this is being run on FILES of previous twitter data
+# with (time, hashtag) format
+# that's why we use the timestamp to calculate time
+# instead of actual time as we would if it were real-time
 class SmartTrendPredictor:
+	# time_str is in format hh:mm:ss
+	# time_str1 comes before time_str2 in time
+	# HAVE NOT HANDLED CASE WHERE WE CROSS DAYS
+	# OR HOURS
+	# OR MINUTES
+	def process_time_diff(time_str1, time_str2):
+		times1 = time_str1.split(':')
+		times2 = time_str2.split(':')
+		del_s = int(times2[2]) - int(times1[2])
+		del_min = int(times2[1]) - int(times1[1])
+		del_hr = int(times2[0]) - int(times1[0])
+		return (del_hr, del_min, del_s)
 	# update the hashtag frequency dict and heap data structures
 	# new_data is in the format (hashtag, timestamp)
-	def update_datastructs(new_data):
+	def update_datastructs(self, new_data):
 		hashtag, timestamp = new_data
+		# update the history
+		# check timestamp and stuff to build current datablocks
+		# diff should be an integer in seconds, maybe
+		diff = process_time_diff(self.end_of_last_block, timestamp)
+		# aggregate only for every block timeunit
+		if diff >= self.threshold:
+			self.hist.aggregate_unit(self.data_block)
+			self.data_block = []
+			self.end_of_last_block = timestamp
+		else: #otherwise, update
+			self.data_block.append(hashtag)
+
+
 		if hashtag in self.hashtag_freq:
 			# update the frequency
 			curr_freq = self.hashtag_freq[hashtag][0]
@@ -30,11 +58,11 @@ class SmartTrendPredictor:
 			# get the ptr and update the heap 
 			ptr = self.hashtag_freq[hashtag][1]
 			entry = self.heap_ptrs[ptr]
-			new_priority = (curr_freq + 0.0)/ self.hokusai[hashtag]
+			new_priority = (curr_freq + 0.0)/ self.hist.query(hashtag)
 			self.heap.decrease_key(entry, new_priority)
 		else:
 			# build the ptr to the heap
-			priority = 1./self.hokusai[hashtag]
+			priority = 1./self.hist.query(hashtag)
 			ptr_val = self.heap.enqueue(hashtag, priority)
 			self.heap_ptrs[self.curr_index] = ptr_val
 			# build hashtag_freq entry
[…]

### 2015-01-11 tweet_trender.py
added some timing stuff

@@ -21,17 +21,19 @@ import key_functions
 # instead of actual time as we would if it were real-time
 class SmartTrendPredictor:
 	# time_str is in format hh:mm:ss
-	# time_str1 comes before time_str2 in time
-	# HAVE NOT HANDLED CASE WHERE WE CROSS DAYS
-	# OR HOURS
-	# OR MINUTES
+	# IMPORTANT: time_str1 < time_str2 in time
 	def process_time_diff(time_str1, time_str2):
 		times1 = time_str1.split(':')
 		times2 = time_str2.split(':')
-		del_s = int(times2[2]) - int(times1[2])
-		del_min = int(times2[1]) - int(times1[1])
-		del_hr = int(times2[0]) - int(times1[0])
-		return (del_hr, del_min, del_s)
+		hr1, hr2, m1, m2, s1, s2 = times1[0], times2[0], times1[1], times2[1], times1[2], times2[2]
+		# in seconds
+		time_diff = 0
+		if hr1 > hr2: # only possible between 11 pm and 12 am, since we're on 24 hour clock
+			hr2 += 24
+		timetot1 = hr1*60*60 + m1*60 + s1
+		timetot2 = hr2*60*60 + m2*60 + s2
+		return timetot2 - timetot1
+
 	# update the hashtag frequency dict and heap data structures
 	# new_data is in the format (hashtag, timestamp)
 	def update_datastructs(self, new_data):

### 2015-01-11 tweet_trender.py
minor fixes

@@ -129,6 +129,7 @@ class SmartTrendPredictor:
 	                tweet_dt = parser.parse(timestamp)
 	                if isFirstLine:
 	                	self.end_of_last_block = timestamp
+	                	isFirstLine = False
 	            except Exception, e:
 	                print e
 	                continue

### 2015-01-11 tweet_trender.py
few fixes

@@ -16,7 +16,7 @@ import key_functions
 
 
 # this is being run on FILES of previous twitter data
-# with (time, hashtag) format
+# with (hashtag, timestamp) format
 # that's why we use the timestamp to calculate time
 # instead of actual time as we would if it were real-time
 class SmartTrendPredictor:
@@ -130,6 +130,7 @@ class SmartTrendPredictor:
 	                if isFirstLine:
 	                	self.end_of_last_block = timestamp
 	                	isFirstLine = False
+	               		update_datastructs((hashtag, timestamp))
 	            except Exception, e:
 	                print e
 	                continue

### 2015-01-11 tweet_trender.py
more updates

@@ -43,34 +43,43 @@ class SmartTrendPredictor:
 		# diff should be an integer in seconds, maybe
 		diff = process_time_diff(self.end_of_last_block, timestamp)
 		# aggregate only for every block timeunit
-		if diff >= self.threshold:
-			self.hist.aggregate_unit(self.data_block)
-			self.data_block = []
-			self.end_of_last_block = timestamp
-		else: #otherwise, update
-			self.data_block.append(hashtag)
 
+		if diff < self.threshold: #otherwise, update
+			self.data_block.append(hashtag)
+			# somehow need to update A here? (self.present)
+			# change paradigm from block 
 
 		if hashtag in self.hashtag_freq:
 			# update the frequency
 			curr_freq = self.hashtag_freq[hashtag][0]
 			curr_freq += 1
-			self.hashtag_freq[hashtag][0] = curr_freq
+			self.hashtag_freq[hashtag][0] = curr_freq	
 
 			# get the ptr and update the heap 
 			ptr = self.hashtag_freq[hashtag][1]
 			entry = self.heap_ptrs[ptr]
-			new_priority = (curr_freq + 0.0)/ self.hist.query(hashtag)
+			# negative since we're using a min heap	
+
+			# NEED TO AVOID INFINITY HERE
+			new_priority = (-1)*(curr_freq + 0.0)/ self.hist.query(hashtag)
 			self.heap.decrease_key(entry, new_priority)
 		else:
 			# build the ptr to the heap
-			priority = 1./self.hist.query(hashtag)
+			# negative since we have a min heap
+			# NEED TO AVOID INFINITY HERE
+			priority = (-1)*1./self.hist.query(hashtag)
 			ptr_val = self.heap.enqueue(hashtag, priority)
 			self.heap_ptrs[self.curr_index] = ptr_val
 			# build hashtag_freq entry
 			self.hashtag_freq[hashtag] = [1, self.curr_index]
 			# new pointer
 			self.curr_index += 1
+
+		if diff >= self.threshold:
+			self.hist.aggregate_unit(self.data_block)
+			self.data_block = []
+			self.end_of_last_block = timestamp
+			# update all other structs
 
 	def __init__(self, data_file):
 
@@ -136,6 +145,20 @@ class SmartTrendPredictor:
 	                continue
 	            self.tweets[hashtag.lower()].append(tweet_dt)
 	            self.inverted_tweets_idx[tweet_dt].append(hashtag.lower())
+
+	def get_topk_hashtags(self, k):
[…]

### 2015-01-11 main.py
Add naive novelty detection

@@ -4,5 +4,7 @@ from naive_trend_predictor import NaiveTrendPredictor
 
 ntp = NaiveTrendPredictor('tweet_data')
 eastern = pytz.timezone('US/Eastern')
-print ntp.get_trending(5, eastern.localize(datetime(2014, 12, 30, 16, 45)),
+print ntp.get_most_popular(5, eastern.localize(datetime(2014, 12, 30, 16, 45)),
+                       eastern.localize(datetime(2014, 12, 30, 17, 0)))
+print ntp.get_most_novel(5, eastern.localize(datetime(2014, 12, 30, 16, 45)),
                        eastern.localize(datetime(2014, 12, 30, 17, 0)))

### 2015-01-11 naive_trend_predictor.py

@@ -18,11 +18,40 @@ class NaiveTrendPredictor:
                 self.tweets[hashtag.lower()].append(tweet_dt)
                 self.inverted_tweets_idx[tweet_dt].append(hashtag.lower())
 
-    def get_trending(self, n, start_dt, end_dt):
+
+    def get_hashtags(self, start_dt, end_dt):
         hashtag_grps = (self.inverted_tweets_idx[tweet_dt]
                         for tweet_dt in self.inverted_tweets_idx
                         if start_dt <= tweet_dt and tweet_dt <= end_dt)
-        most_popular = {(sum((start_dt <= tweet_dt and tweet_dt <= end_dt) for tweet_dt in self.tweets[hashtag]), hashtag)
-                        for hashtag_grp in hashtag_grps
-                        for hashtag in hashtag_grp}
+        hashtags = (hashtag for hashtag_grp in hashtag_grps
+                    for hashtag in hashtag_grp)
+        return hashtags
+
+
+    def get_hashtag_freq(self, hashtag, start_dt, end_dt):
+        return sum(start_dt <= tweet_dt and tweet_dt <= end_dt
+                   for tweet_dt in self.tweets[hashtag])
+
+
+    def get_most_popular(self, n, start_dt, end_dt):
+        hashtags = self.get_hashtags(start_dt, end_dt)
+        most_popular = {(self.get_hashtag_freq(hashtag, start_dt, end_dt), hashtag)
+                        for hashtag in hashtags}
         return list(itertools.islice(reversed(sorted(most_popular)), 0, n))
+
+
+    def get_most_novel(self, n, start_dt, end_dt):
+        prev_start_dt = start_dt - (end_dt - start_dt)
+
+        prev_hashtags = self.get_hashtags(prev_start_dt, start_dt)
+        cur_hashtags = self.get_hashtags(start_dt, end_dt)
+
+        prev_freq_counts = {hashtag: self.get_hashtag_freq(hashtag, prev_start_dt, start_dt)
+                            for hashtag in prev_hashtags}
+        cur_freq_counts = {hashtag: self.get_hashtag_freq(hashtag, start_dt, end_dt)
+                           for hashtag in cur_hashtags}
+
+        most_novel = (((cur_freq_counts[hashtag] + 1.0)/(prev_freq_counts.get(hashtag, 0) + len(prev_freq_counts)), hashtag)
+                      for hashtag in cur_freq_counts)
+
+        return list(itertools.islice(reversed(sorted(most_novel)), 0, n))

### 2015-01-11 naive_trend_predictor.py
Fix smoothing amount

@@ -2,6 +2,8 @@ from collections import defaultdict
 import itertools
 from dateutil import parser
 
+TRENDING_THRESHOLD = 50
+
 class NaiveTrendPredictor:
     def __init__(self, data_file):
         self.tweets = defaultdict(list)
@@ -51,7 +53,8 @@ class NaiveTrendPredictor:
         cur_freq_counts = {hashtag: self.get_hashtag_freq(hashtag, start_dt, end_dt)
                            for hashtag in cur_hashtags}
 
-        most_novel = (((cur_freq_counts[hashtag] + 1.0)/(prev_freq_counts.get(hashtag, 0) + len(prev_freq_counts)), hashtag)
+        most_novel = (((cur_freq_counts[hashtag] + 1.0)/
+                       (prev_freq_counts.get(hashtag, 0) + TRENDING_THRESHOLD), hashtag)
                       for hashtag in cur_freq_counts)
 
         return list(itertools.islice(reversed(sorted(most_novel)), 0, n))

### 2015-01-11 history.py
history is finished being implemented, tested and debugged

@@ -80,6 +80,10 @@ class History(object):
 		# present is a count-min sketch containing
 		# sub-unit time counts of indexes. 
 		self.present = CMSketch(m, d)
+		# ready is a t/f value to determine whether or not
+		# to use the present as a score while the aggregate weighted score
+		# is being computed
+		self.ready = False
 		# use a CM-sketch to keep track of aggregate weighted score
 		# A = sum{j = 1 to log T} (M^j / 2^j)
 		# (we add the present ourselves)
@@ -124,6 +128,7 @@ class History(object):
 			# update present as we update the accumulator
 			self.present.add(data, 1)
 
+		self.ready = False
 		# we update the whole structure with M_bar
 		# we calculate l: 
 		# l = max over all i such that (t mod 2^i) == 0
@@ -140,26 +145,29 @@ class History(object):
 		# go up to the index that is find_l + 1, or the max index
 		# if find_l + 1 >= to it
 		for i in range(min(find_l(self.t) + 1, self.n)):
+			# now we want to add the appropriate value: A + 1/2^(i)(M_bar - M^j)
+			# M_bar - M^j
+			difference = sketch_sum(accumulator, sketch_scalar_product(self.cm_sketch_list[i], -1))
+			# A = A + (1/2)^i difference
+			self.aggregate_score = sketch_sum(self.aggregate_score, 
+									sketch_scalar_product(difference, pow(0.5, i)))
 			# temporary storage
 			T = deepcopy(accumulator)
 			# aggregate into accumulator for next round
 			accumulator = sketch_sum(accumulator, self.cm_sketch_list[i])
 			# set the value
 			self.cm_sketch_list[i] = T
-			# now we want to add the appropriate value of T multiplied by a factor of 1/2^(i + 1)
-			self.aggregate_score = sketch_sum(self.aggregate_score, sketch_scalar_product(T, pow(0.5, (i + 1))))
+		# now we're ready to use CM-sketch values
+		self.ready = True
 		# reset the present now that we're done with one time block
 		self.present = CMSketch(self.m, self.d)
 
-
-	# CURRENTLY THERE IS A MISMATCH BETWEEN THE TWO-- BUG
-	
 	# we want to put these values into its own count-min sketch, (call it A)
 	# updated in sync so as to not waste log T time summing
 	# for each query.
 	# this value will provide a key for our heap
 	def query_slow(self, x):
[…]

### 2015-01-11 naive_trend_predictor.py
Preprocess duplicate hashtags

@@ -2,7 +2,7 @@ from collections import defaultdict
 import itertools
 from dateutil import parser
 
-TRENDING_THRESHOLD = 50
+TRENDING_THRESHOLD = 25
 
 class NaiveTrendPredictor:
     def __init__(self, data_file):
@@ -31,12 +31,12 @@ class NaiveTrendPredictor:
 
 
     def get_hashtag_freq(self, hashtag, start_dt, end_dt):
-        return sum(start_dt <= tweet_dt and tweet_dt <= end_dt
-                   for tweet_dt in self.tweets[hashtag])
+        return sum(1 for tweet_dt in self.tweets[hashtag]
+                   if start_dt <= tweet_dt and tweet_dt <= end_dt)
 
 
     def get_most_popular(self, n, start_dt, end_dt):
-        hashtags = self.get_hashtags(start_dt, end_dt)
+        hashtags = set(self.get_hashtags(start_dt, end_dt))
         most_popular = {(self.get_hashtag_freq(hashtag, start_dt, end_dt), hashtag)
                         for hashtag in hashtags}
         return list(itertools.islice(reversed(sorted(most_popular)), 0, n))
@@ -45,8 +45,8 @@ class NaiveTrendPredictor:
     def get_most_novel(self, n, start_dt, end_dt):
         prev_start_dt = start_dt - (end_dt - start_dt)
 
-        prev_hashtags = self.get_hashtags(prev_start_dt, start_dt)
-        cur_hashtags = self.get_hashtags(start_dt, end_dt)
+        prev_hashtags = set(self.get_hashtags(prev_start_dt, start_dt))
+        cur_hashtags = set(self.get_hashtags(start_dt, end_dt))
 
         prev_freq_counts = {hashtag: self.get_hashtag_freq(hashtag, prev_start_dt, start_dt)
                             for hashtag in prev_hashtags}

### 2015-01-11 history.py
updated tweet_trender.py

@@ -98,6 +98,12 @@ class History(object):
 		for i in range(n):
 			self.cm_sketch_list.append(CMSketch(m, d))
 
+	def update_present_only(self, datum):
+		self.ready = False
+		# don't update the full time
+		# this is a sub-unit update
+		self.present.add(datum, 1)
+
 	# data_block is a block of data, presented as an iterable object
 	# the block of data consists of data that arrived in a single time unit
 	# implements algorithm 2 from the paper
@@ -123,6 +129,10 @@ class History(object):
 		# data structure with the exact frequencies that we can
 		# query for exact frequencies. 
 		# with frequency 1 for each appearance
+
+		# reset the present when we aggregate the whole thing
+		self.present = CMSketch(self.m, self.d)
+		# (data_block is the present)
 		for data in data_block:
 			accumulator.add(data, 1)
 			# update present as we update the accumulator

### 2015-01-11 tweet_trender.py

@@ -5,6 +5,7 @@ import fib_heap as fh
 from collections import default_dict
 import history as h
 import key_functions
+from dateutil import parser
 
 # when parsing tweets, need to make sure 
 # the tweet sending model (from file)
@@ -20,6 +21,7 @@ import key_functions
 # that's why we use the timestamp to calculate time
 # instead of actual time as we would if it were real-time
 class SmartTrendPredictor:
+	'''
 	# time_str is in format hh:mm:ss
 	# IMPORTANT: time_str1 < time_str2 in time
 	def process_time_diff(time_str1, time_str2):
@@ -33,21 +35,31 @@ class SmartTrendPredictor:
 		timetot1 = hr1*60*60 + m1*60 + s1
 		timetot2 = hr2*60*60 + m2*60 + s2
 		return timetot2 - timetot1
+	'''
 
 	# update the hashtag frequency dict and heap data structures
 	# new_data is in the format (hashtag, timestamp)
 	def update_datastructs(self, new_data):
-		hashtag, timestamp = new_data
+		hashtag, curr_dt = new_data
+		#curr_dt = parser.parse(timestamp)
 		# update the history
 		# check timestamp and stuff to build current datablocks
 		# diff should be an integer in seconds, maybe
-		diff = process_time_diff(self.end_of_last_block, timestamp)
-		# aggregate only for every block timeunit
+		diff = curr_dt - self.end_of_last_block
 
-		if diff < self.threshold: #otherwise, update
+		# aggregate only for every block timeunit
+		if diff > self.block_threshold:
+			self.hist.aggregate_unit(self.data_block)
+			self.data_block = []
+			self.end_of_last_block = curr_dt
+			# update all other structs
+		else:
 			self.data_block.append(hashtag)
-			# somehow need to update A here? (self.present)
-			# change paradigm from block 
+			# update the present structure in hist
+			# don't worry, this will get cleared when
+			# we add the whole block, and get re-added from the start
+			self.hist.update_present_only(hashtag)
+
 
 		if hashtag in self.hashtag_freq:
 			# update the frequency
@@ -60,13 +72,13 @@ class SmartTrendPredictor:
 			entry = self.heap_ptrs[ptr]
 			# negative since we're using a min heap	
 
-			# NEED TO AVOID INFINITY HERE
+			# query value should never be 0
 			new_priority = (-1)*(curr_freq + 0.0)/ self.hist.query(hashtag)
 			self.heap.decrease_key(entry, new_priority)
[…]

### 2015-01-11 tweet_trender.py
finishing up tweet trender

@@ -3,6 +3,7 @@
 
 import fib_heap as fh
 from collections import default_dict
+from collections import deque
 import history as h
 import key_functions
 from dateutil import parser
@@ -52,7 +53,7 @@ class SmartTrendPredictor:
 			self.hist.aggregate_unit(self.data_block)
 			self.data_block = []
 			self.end_of_last_block = curr_dt
-			# update all other structs
+
 		else:
 			self.data_block.append(hashtag)
 			# update the present structure in hist
@@ -60,7 +61,7 @@ class SmartTrendPredictor:
 			# we add the whole block, and get re-added from the start
 			self.hist.update_present_only(hashtag)
 
-
+		# update all other structs
 		if hashtag in self.hashtag_freq:
 			# update the frequency
 			curr_freq = self.hashtag_freq[hashtag][0]
@@ -87,6 +88,36 @@ class SmartTrendPredictor:
 			# new pointer
 			self.curr_index += 1
 
+		# update the queue
+		# (appending to the right, most recent)
+		self.input_queue.append(new_data)
+
+		# check first thing in the queue (left)
+		# while the time of the element of the
+		# queue + y is less than current time
+		# (that means it's too old)
+		# remove it from the "current" data structures
+		# (the heap, the freq hashtable, the queue)
+		# (since we delete from the left, the index to look at is always 0)
+		while self.input_queue[0][1] + self.y_res < curr_dt:
+			hashtag_old, old_dt = self.input_queue.popleft()
+			new_freq = self.hashtag_freq[hashtag_old][0] - 1
+			if new_freq > 0:
+				# update hashtag frequency table
+				self.hashtag_freq[hashtag_old][0] = new_freq
+				# update heap 
+				ptr = self.hashtag_freq[hashtag_old][1]
+				entry = self.heap_ptrs[ptr]
+				new_priority = (-1) * (new_freq + 0.0) / self.hist.query(hashtag_old)
+				self.heap.decrease_key(entry, new_priority)
+			else:
+				ptr = self.hashtag_freq[hashtag_old][1]
+				# delete the pointer to this entry from the heap, since its frequency is 0
+				self.heap.delete(ptr)
+				# delete the entry from the hashtable
+				del self.hashtag_freq[hashtag_old]
+
+
 
 	def __init__(self, data_file):
 
@@ -98,18 +129,26 @@ class SmartTrendPredictor:
 
 		# number of seconds in a day
 		# a day is a block, can modify this
+		# this is for the history datastructure
 		self.block_threshold = 60*60*24
[…]

### 2015-01-12 tweet_trender.py
finished tweet_trender

@@ -2,11 +2,13 @@
 # Algorithm Implementation 
 
 import fib_heap as fh
-from collections import default_dict
+from collections import defaultdict
 from collections import deque
 import history as h
 import key_functions
 from dateutil import parser
+from datetime import timedelta
+from copy import deepcopy
 
 # when parsing tweets, need to make sure 
 # the tweet sending model (from file)
@@ -21,103 +23,17 @@ from dateutil import parser
 # with (hashtag, timestamp) format
 # that's why we use the timestamp to calculate time
 # instead of actual time as we would if it were real-time
-class SmartTrendPredictor:
-	'''
-	# time_str is in format hh:mm:ss
-	# IMPORTANT: time_str1 < time_str2 in time
-	def process_time_diff(time_str1, time_str2):
-		times1 = time_str1.split(':')
-		times2 = time_str2.split(':')
-		hr1, hr2, m1, m2, s1, s2 = times1[0], times2[0], times1[1], times2[1], times1[2], times2[2]
-		# in seconds
-		time_diff = 0
-		if hr1 > hr2: # only possible between 11 pm and 12 am, since we're on 24 hour clock
-			hr2 += 24
-		timetot1 = hr1*60*60 + m1*60 + s1
-		timetot2 = hr2*60*60 + m2*60 + s2
-		return timetot2 - timetot1
-	'''
-
-	# update the hashtag frequency dict and heap data structures
-	# new_data is in the format (hashtag, timestamp)
-	def update_datastructs(self, new_data):
-		hashtag, curr_dt = new_data
-		#curr_dt = parser.parse(timestamp)
-		# update the history
-		# check timestamp and stuff to build current datablocks
-		# diff should be an integer in seconds, maybe
-		diff = curr_dt - self.end_of_last_block
-
-		# aggregate only for every block timeunit
-		if diff > self.block_threshold:
-			self.hist.aggregate_unit(self.data_block)
-			self.data_block = []
-			self.end_of_last_block = curr_dt
-
-		else:
-			self.data_block.append(hashtag)
-			# update the present structure in hist
-			# don't worry, this will get cleared when
-			# we add the whole block, and get re-added from the start
-			self.hist.update_present_only(hashtag)
-
-		# update all other structs
-		if hashtag in self.hashtag_freq:
-			# update the frequency
-			curr_freq = self.hashtag_freq[hashtag][0]
-			curr_freq += 1
-			self.hashtag_freq[hashtag][0] = curr_freq	
[…]

### 2015-01-12 tweet_trender.py
commented tweet_trender usage a bit more

@@ -32,9 +32,15 @@ from copy import deepcopy
 # the assumption is that the tweets are in chronological order.
 # the trend predictor builds a datastructure that evalutes 
 # the top k tweets for that specific file. 
-# we want to implement time snapshots
+# maybe we want to implement time snapshots INSIDE a given file?
+# as in, top k for a certain time range? maybe not necessary, 
+# because here, we're just simulating "up to a certain point in time"
+# with the file 
 class SmartTrendPredictor(object):
 
+	# NOTE THE HISTORY PARAMETERS, THRESHOLDS, AND OTHER STUFF
+	# MAY BE CHANGED INSIDE HERE. 
+	# I COULD EXPOSE THESE PARAMETERS BUT I DID NOT. feel free to do so if you want.
 	def __init__(self, data_file):
 
 		# maintains list of hashtags
@@ -246,6 +252,10 @@ class SmartTrendPredictor(object):
 			print 'count of ' + str(key) + ': ' + str(key_counts[key])
 
 
+# build a smart trend predictor, 
+# see what the key counts are (what priorities exist),
+# see what the top ones are for the test file 'tweet_data', 
+# a file that only spans a few hours (so the history stuff doesn't get tested..)
 def test():
 	stp = SmartTrendPredictor('tweet_data')
 	stp.print_all_keys()

### 2015-01-12 tweet_trender.py
fixed pointer update problem in get_topk_hashtags

@@ -225,6 +225,8 @@ class SmartTrendPredictor(object):
 
 	# the top hashtags in the min-heap
 	# (most negative)
+	# FORGOT TO CHECK POINTERS HERE
+	# need to write code that removes properly
 	def get_topk_hashtags(self, k):
 		# what we will return
 		# list of entries (have priority and value)
@@ -233,11 +235,16 @@ class SmartTrendPredictor(object):
 			top_node = self.heap.dequeue_min()
 			val = top_node.get_value()
 			priority = top_node.get_priority()
+			ptr = self.hashtag_freq[val][1]
+			# self.heap_ptrs[ptr] should be the same as top_node
+
 			#print 'Priority: ' + str(-1*priority)
 			#print 'Value: ' + str(val)
-			best_nodes.append((priority, val))
+			best_nodes.append((priority, val, ptr))
 		for node in best_nodes:
-			self.heap.enqueue(node[1], node[0])
+			new_node = self.heap.enqueue(node[1], node[0])
+			ptr = node[2]
+			self.heap_ptrs[ptr] = new_node
 		return sorted(best_nodes, reverse=True)
 
 	def print_all_keys(self):

### 2015-01-12 history.py
fix tweet_trender

@@ -188,7 +188,7 @@ class History(object):
 		if self.ready:
 			return self.aggregate_score.query(x)
 		else: # only if we're not ready 
-			return self.present.query(x)
+			return self.present.query(x) + self.aggregate_score.query(x)
 
 
 # test the history functions

### 2015-01-12 tweet_trender.py

@@ -129,6 +129,7 @@ class SmartTrendPredictor(object):
 
 		# aggregate only for every block timeunit
 		if diff > self.block_threshold:
+			self.data_block.append(hashtag)
 			self.hist.aggregate_unit(self.data_block)
 			self.data_block = []
 			self.end_of_last_block = curr_dt

### 2015-01-12 tweet_trender.py
Removing unpacking err

@@ -105,7 +105,9 @@ class SmartTrendPredictor(object):
 		with open(data_file) as f:
 			for line in f:
 				try:
-					id_, timestamp, hashtag = line.strip().split(',')
+                                        splitLine = line.strip().split(',')
+                                        timestamp = splitLine[1]
+                                        hashtag = ','.join(splitLine[2:])
 					tweet_dt = parser.parse(timestamp)
 					if isFirstLine:
 						# 'last blocks' start at the beginning of the file

### 2015-01-12 521_project_paper.tex
changed parameters of the history data struct

--- 2015-01-07 Project writeup.tex
@@
 \section{Introduction}
+
+Things to add:
+abstract
+clear definition of problem and motivation 
+adapt proposal -> introduction stuff
+background work: cm-sketch, hokusai (algorithm), specific twitter trending work: ? 
+http://ieeexplore.ieee.org/xpls/abs_all.jsp?arnumber=6137387&tag=1 
+(this is the closest we could find)
+(to the best of our knowledge, this work has not been done before)
+
+how did we approach it 
+overview
+
+data structures
+
+algorithms
+clean up algorithm section
+- make sure changes in implementation are reflected in algorithm
+
+justification
+
+results
+
+naive algorithm serves as comparison
+description of naive
+
+the results
+
+future work
+
+references
 

### 2015-01-12 tweet_trender.py

@@ -93,8 +93,15 @@ class SmartTrendPredictor(object):
 
 		# similar to Hokusai data structure 
 		# pick some parameters here
-		# n = 10, m = 1000 (size of hash tables), d = 100 (# of tables)
-		self.hist = h.History(10, 1000, 100)
+		# n = 6, m = 100000 (size of hash tables), d = 20 (# of tables)
+		# n: we do it for a month, so ~ 2^(6-1) days is definitely enough
+		# m = ceiling(e/epsilon), d = ceiling(ln(1/delta))
+		# error in query is in factor epilson with probability 1 - delta
+		# for d = 20, we have with high probability (1 - 1/e^20)
+		# we have error factor of e/100000 of the total number of hashtags
+		# total number of hashtags is ~ 15 million
+		# thus error bounds are in range ~ 450 on a given count
+		self.hist = h.History(6, 100000, 20)
 
 		# first line bool
 		# used for checking if its the first line in the file
@@ -105,9 +112,9 @@ class SmartTrendPredictor(object):
 		with open(data_file) as f:
 			for line in f:
 				try:
-                                        splitLine = line.strip().split(',')
-                                        timestamp = splitLine[1]
-                                        hashtag = ','.join(splitLine[2:])
+					splitLine = line.strip().split(',')
+					timestamp = splitLine[1]
+					hashtag = ','.join(splitLine[2:])
 					tweet_dt = parser.parse(timestamp)
 					if isFirstLine:
 						# 'last blocks' start at the beginning of the file

### 2015-01-12 521_project_paper.tex
added outline

@@
 
+to note: our goal was to make a real time query algorithm that USES
+the history (in condensed form) to help understand current time 
+we're remembering the past with less accuracy to save space (realistic model)
+in order to get good estimates of present trending
+
+this means we are not trying to query the past the way Hokusai is.
+
 how did we approach it 
@@
 data structures
+description
+-- fib heap
+-- queue
+-- frequency hash table
+-- History (own subsection)
 
 algorithms
-clean up algorithm section
+description
 - make sure changes in implementation are reflected in algorithm 
+-- querying the (History)
+-- updating the data structure
+------ adding to the data structure
+----------- frequency hash table update/ heap update / (update history every time block - mention why we can add, etc, some of the theory from the Hokusai paper)
+------ removing from the data structure (our snapshot of the present)
+----------- every time unit of the present, we remove outdated frequencies
+-- top k hashtags (real time)
+
+runtime analysis
+space analysis
+accuracy analysis
+
+design choices for data structure
+-- data structure modification (History vs Hokusai)
+-- fibonacci heap 
+-- heap key function
 

### 2015-01-12 521_project_paper.tex
Updated algorithms and analysis

@@
 background work: cm-sketch, hokusai (algorithm), specific twitter trending work: ? 
-http://ieeexplore.ieee.org/xpls/abs_all.jsp?arnumber=6137387&tag=1 
+%http://ieeexplore.ieee.org/xpls/abs_all.jsp?arnumber=6137387&tag=1 
 (this is the closest we could find)
@@
 To this structure we add another layer that combines these Count-Min sketches into a single Count-Min sketch that depreciates the value of older data.
-This aggregate Count-Min sketch stores the values $\bar{M} + \sum\limits_{j=1}^{\log⁡(T)} \frac{M^j}{2^j}$ where $\bar{M}$ is the Count-Min sketch storing today's data, $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ resolution, and $T$ is an upper bound on the number of days stored by the data structure.
+This aggregate Count-Min sketch stores the values $\bar{M} + \sum\limits_{j=0}^{\log⁡(T)} \frac{M^j}{2^j}$ where $\bar{M}$ is the Count-Min sketch storing today's data, $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ resolution, and $T$ is an upper bound on the number of days stored by the data structure.
 In section \ref{Correct} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i$ days ago with value approximately $\frac{1}{2i}$.
@@
 \While {data arrives}
-	\If {hashtag is in hash table}
-		\State Increment the frequency stored at that entry.
-		\State Update the key of the node in the Fibonacci heap.
-	\Else
-		\State Insert a new node into the Fibonacci heap with the appropriate key and value. 
-		\State Insert the hashtag into the hash table with a pointer to this node.
+	\If {the current day is different than that of the last hashtag seen}
+		\State Do all the end-of-day aggregation for the Hokusai structure as detailed in Algorithm \ref{alg1}.
+		\ForAll {elements in the Fibonacci heap}
+			\State look up the hashtag corresponding to this node in the hash table
+			\State Update the key of the node to: \[\frac{\text{frequency in last y minutes found at the table entry}}{\text{new value in Hokusai data structure}}\]
+		\EndFor
 	\EndIf
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Thread to discard oldest recent hashtags.}
-\begin{algorithmic}[1]
-\Loop
 	\If {queue is not empty}
@@
 	\EndIf
-\EndLoop
-\State Make the hash table for the oldest $x$ minute interval correspond to the upcoming $x$ minute interval.
-
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Updates at end of a day.}
-\begin{algorithmic}[1]
-\State Do all the end-of-day aggregation for the Hokusai structure as detailed in Algorithm \ref{alg1}.
-\ForAll {elements in the Fibonacci heap}
-	\State look up the hashtag corresponding to this node in the hash table
-	\State Update the key of the node to: \[\frac{\text{frequency in last y minutes found at the table entry}}{\text{new value in Hokusai data structure}}\]
-\EndFor
+	\If {hashtag is in hash table}
+		\State Increment the frequency stored at that entry.
+		\State Update the key of the node in the Fibonacci heap.
+	\Else
+		\State Insert a new node into the Fibonacci heap with the appropriate key and value. 
+		\State Insert the hashtag into the hash table with a pointer to this node.
+ 	\EndIf
+\EndWhile
 \end{algorithmic}
@@
 Takes O(m) time to compute each key.
-Takes O(1) amortized time to update each key since all keys are nondecreasing.
+Takes O(log(s)) amortized time to update each key.
 
-Thus, takes O(nm + s) amortized time + O(ms) time at the end of the day.
+Thus, takes O(nm + slog(s)) amortized time + O(ms) time at the end of the day.
 

### 2015-01-12 tweet_trender.py
normalized hashtag, reverted to different history parameters

@@ -93,15 +93,16 @@ class SmartTrendPredictor(object):
 
 		# similar to Hokusai data structure 
 		# pick some parameters here
-		# n = 6, m = 100000 (size of hash tables), d = 20 (# of tables)
+		# n = 6, m = 3500 (size of hash tables), d = 20 (# of tables)
 		# n: we do it for a month, so ~ 2^(6-1) days is definitely enough
 		# m = ceiling(e/epsilon), d = ceiling(ln(1/delta))
 		# error in query is in factor epilson with probability 1 - delta
 		# for d = 20, we have with high probability (1 - 1/e^20)
-		# we have error factor of e/100000 of the total number of hashtags
+		# we have error factor of e/3500 of the total number of hashtags
 		# total number of hashtags is ~ 15 million
-		# thus error bounds are in range ~ 450 on a given count
-		self.hist = h.History(6, 100000, 20)
+		# thus error bounds are in range ~ 12000 on a given count
+		# this error provides a 'good' amount of smoothing
+		self.hist = h.History(6, 3500, 20)
 
 		# first line bool
 		# used for checking if its the first line in the file
@@ -115,6 +116,8 @@ class SmartTrendPredictor(object):
 					splitLine = line.strip().split(',')
 					timestamp = splitLine[1]
 					hashtag = ','.join(splitLine[2:])
+					# normalize hashtags
+					hashtag = hashtag.lower()
 					tweet_dt = parser.parse(timestamp)
 					if isFirstLine:
 						# 'last blocks' start at the beginning of the file

### 2015-01-12 521_project_paper.tex
Changes to paper

@@
 \usepackage{amsthm}
+\usepackage{booktabs}
 
@@
 This aggregate Count-Min sketch stores the values $\bar{M} + \sum\limits_{j=0}^{\log⁡(T)} \frac{M^j}{2^j}$ where $\bar{M}$ is the Count-Min sketch storing today's data, $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ resolution, and $T$ is an upper bound on the number of days stored by the data structure.
-In section \ref{Correct} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i$ days ago with value approximately $\frac{1}{2i}$.
+In section \ref{Correct} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i > 0$ days ago with value approximately $\frac{1}{i}$.
 
@@
 \EndFor
-\State Initialize $t = 0$, $\bar{M} = 0$, and $A = 0$
+\State Initialize $t = 0$
+\State Initialize Count-Min sketches $\bar{M} = 0$ and $A = 0$
 \While {data arrives}
@@
 
-Maybe put table here of all tunable parameters mentioned in the paper to remind readers what they refer to.
-
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[h]
+\caption{Variables referenced in this section}
+\begin{tabular}{@{}ll@{}}
+\toprule
+Symbol & Meaning                                                           \\ \midrule
+m      & Number of tables for each Count-Min sketch                        \\
+n      & Size of each hash table in each Count-Min sketch                  \\
+s      & Number of distinct hashtags in the Current Window                 \\
+T      & Upper bound on the number of days worth of data stored in History \\
+y      & Minutes of data contained in the Current Window                   \\ \bottomrule
+\end{tabular}
+\end{table}
 \subsubsection{Time analysis}
-
-Let s = size of heap = number of distinct hashtags seen in last y minutes.
 

### 2015-01-13 viz/trending.js
Add visualization

@@ -1,0 +1,41 @@
+//var allData = [[0, 9, 8, 0, 0, 0, 11, 26, 0, 1, 0, 3, 0, 44, 5, 0, 0, 5, 1, 15, 0, 0, 1, 1, 0, 1, 7, 26, 0, 0, 2, 6, 1, 0, 0, 0, 0, 0, 558, 63, 0, 0, 5], [0, 29, 169, 2, 0, 0, 184, 35, 0, 0, 0, 3, 0, 153, 7, 0, 0, 4, 3, 17, 0, 0, 3, 10, 0, 1, 10, 46, 0, 0, 12, 14, 3, 0, 3, 1, 2, 0, 469, 27, 0, 0, 2], [0, 44, 302, 0, 1, 349, 311, 26, 0, 5, 0, 2, 0, 290, 4, 1, 0, 5, 3, 141, 0, 0, 1, 3, 0, 5, 17, 60, 0, 0, 6, 5, 8, 0, 0, 0, 5, 0, 16, 44, 0, 2, 17], [0, 67, 89, 0, 0, 282, 87, 24, 0, 4, 0, 3, 0, 305, 3, 2, 2, 7, 1, 583, 0, 0, 0, 4, 0, 7, 15, 58, 0, 0, 5, 13, 4, 0, 12, 0, 4, 1, 109, 38, 1, 0, 6], [0, 137, 73, 0, 0, 17, 71, 26, 0, 4, 0, 5, 0, 353, 3, 0, 2, 12, 1, 122, 0, 0, 0, 5, 0, 2, 35, 39, 0, 0, 4, 4, 18, 0, 7, 0, 7, 0, 44, 14, 1, 1, 10], [0, 1389, 207, 0, 1, 0, 216, 12, 0, 0, 0, 2, 0, 483, 2, 1, 1, 17, 1, 23, 0, 0, 2, 4, 0, 0, 38, 51, 0, 0, 5, 12, 656, 0, 2, 0, 5, 0, 5, 13, 0, 0, 7], [0, 547, 187, 0, 0, 1, 192, 19, 0, 1, 0, 1, 0, 208, 1, 0, 0, 45, 1, 14, 0, 0, 0, 4, 0, 2, 43, 36, 0, 0, 2, 4, 180, 0, 0, 0, 4, 0, 2, 8, 0, 0, 6], [0, 78, 202, 0, 1, 6, 210, 13, 0, 0, 0, 1, 0, 23, 4,
[…]

### 2015-01-13 521_project_paper.tex

@@
 To store data about all occurrences of every hashtag seen in the past, we use a modified version of the time-aggregated Hokusai system\footnote{http://arxiv.org/pdf/1210.4891v1.pdf}, which is an extension of the Count-Min sketch.
-This data structure is described in full detail in section \ref{Hokusai}.
+This data structure, which we refer to as the History, is described in full detail in section \ref{Hokusai}.
 
@@
 Our data structure for holding hashtags seen in the last $y$ minutes consists of three components: a max Fibonacci heap, a queue, and a hash table.
+We refer to these components collectively as the Current Window.
 
@@
 s      & Number of distinct hashtags in the Current Window                 \\
-T      & Upper bound on the number of days worth of data stored in History \\
+T      & Upper bound on the number of days of data stored in History \\
+x      & Total number of hashtags in the Current Window                   \\
 y      & Minutes of data contained in the Current Window                   \\ \bottomrule
@@
 
-Time for actions when a hashtag comes in:
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[h]
+\caption{Time analysis summary}
+\label{my-label}
+\begin{tabular}{@{}ll@{}}
+\toprule
+Case                                                                                                   & Amortized Time          \\ \midrule
+Processing the insertion of a new hashtag                                                              & $O(m)$                  \\
+\begin{tabular}[c]{@{}l@{}}Processing the removal of a hashtag from \\ the Current Window\end{tabular} & $O(m + \log(s))$        \\
+\begin{tabular}[c]{@{}l@{}}Updating History and Current Window \\ at the end of a day\end{tabular}     & $O(nm + ms + s\log(s))$ \\
+Querying for the top $k$ trending items                                                                & $O(k\log(s))$           \\ \bottomrule
+\end{tabular}
+\end{table}
 
-	Takes amortized O(m) time to update Hokusai structure.
+Processing the insertion of a hashtag takes the following time.  It takes amortized O(m) time to update the History.  It takes expected O(1) time to check if in hash table.
 
-	Takes expected O(1) time to check if in hash table.  If so:
+If it is, it requires O(1) time to increment the frequency, O(m) time to compute the new key, and O(1) amortized time to update the key since it will be nondecreasing.
 
-		O(1) time to increment
+Otherwise, it requires O(m) time to compute the key value, O(1) time to insert the new node in the heap, and O(1) time to insert into the hash table.
 
-		O(m) time to compute key
+Thus, our algorithm takes O(m) amortized time + expected O(1) time to process a new hashtag.
 
-		O(1) amortized time to update key since key nondecreasing
+Processing the removal of a hashtag from the Current Window takes the following time.
+It takes O(1) time to verify that the queue is not empty.
+It takes O(1) time to look at the end of the queue and verify that the timestamp + $y$ minutes is before the current time.
+It takes O(1) time in expectation to look up this hashtag and decrement its frequency.
+Then, it takes O(1) time to check if the frequency is 0.
 
-	Else:
+If so, it takes O(log(s)) amortized time to delete the node in the heap and O(1) time to delete the entry in the hash table.
 
-		O(m) time to compute key value and O(1) time to insert node in heap
+Otherwise, it takes O(m) amortized time to compute the new key for the hash table and O(log(s)) amortized time to update the heap given this key.
 
-		Takes O(1) time to insert into hash table.
+Thus, our algorithm requires O(log(s)) amortized time + expected O(1) time  + O(m) time to remove a hashtag from the Current Window.
 
-Thus, takes O(m) amortized time + expected O(1) time to do the actions when a hashtag comes in.
+The end-of-day updates to the History and the resulting updates to the heap take the following time.
+By Lemma 5 of the Hokusai paper, the amortized time required to do all end-of-day aggregation is O(nm).
+Then, for each of the s nodes in the heap, it takes O(m) time to compute each updated key and O(log(s)) amortized time to update the heap given the new key.
 
-Time for actions when $y$ minutes pass after a hashtag comes in:
-	O(1) time to check if queue is empty.  If not:
-		O(1) time to look at the queue and see if the timestamp + $y$ minutes is before the current time.  If so:
-			Expected O(1) time to look up in hash table and decrement the frequency.
-			O(1) time to check if frequency is 0.  If so:
-				O(log(s)) amortized time to delete node in heap.
-				O(1) time to delete entry in hash table.
-			Else:
-				O(m) amortized time to compute the 
-				O(log(s)) amortized time to update the heap given this key.
+Thus, it takes O(nm + slog(s)) amortized time + O(ms) time to do all necessary updates at the end of the day.
 
-Thus, takes O(log(s)) amortized time + expected O(1) time  + O(m) time to do actions when a hashtag is discarded.
+Querying the data structure for the top k trending items takes O(k*log(s)) amortized time for delete-max operations, O(k) time to announce that these items are trending, and O(k) amortized time to reinsert these nodes into the heap.
 
-Time for actions when day ends:
-Takes O(nm) amortized time to do all end of day aggregation.
-Takes O(m) time to compute each key.
-Takes O(log(s)) amortized time to update each key.
-
-Thus, takes O(nm + slog(s)) amortized time + O(ms) time at the end of the day.
-
-Time for actions when finding top k trending items:
-	Takes O(k*log(s)) time for delete-max operations.
-	Takes O(k) time to announce these as trending.
-Takes O(k) amortized time to reinsert these nodes into the heap.
-
-Thus, takes O(k*log(s)) time + O(k) amortized time to determine what’s trending.
+Thus, it takes O(k*log(s)) amortized time to determine what’s trending.
 
@@
 
-Hokusai structure takes O(nm*log(T)) space.
+The History requires O(nm) space for each Count-Min sketch, so it requires a total of O(nm*log(T)) space.
 
-Heap takes O(s) space.
+Each node in the heap requires a constant amount of space, so the heap requires O(s) space.
 
-Hash table needs O(s) space to keep expected O(1) lookup time.
+The hash table always contains at most s entries with each entry requiring a constant amount of space.
+Also, in order to maintain an expected O(1) lookup time, the hash table needs to have O(s) bins.
+Thus, the hash table requires O(s) space.
 
-Queue requires O(t) space. (where t max hashtags seen in y min interval)
+The queue requires an entry for every hashtag still in the current window, so it requires O(x) space.
 
-Thus, everything requires O(nm*log(T) + s + t) space.
+Thus, everything requires O(nm*log(T) + x) space since s $<$ x.
 

### 2015-01-13 viz/trending.js
Add figures, fix  typos

@@ -1,11 +1,8 @@
-//var allData = [[0, 9, 8, 0, 0, 0, 11, 26, 0, 1, 0, 3, 0, 44, 5, 0, 0, 5, 1, 15, 0, 0, 1, 1, 0, 1, 7, 26, 0, 0, 2, 6, 1, 0, 0, 0, 0, 0, 558, 63, 0, 0, 5], [0, 29, 169, 2, 0, 0, 184, 35, 0, 0, 0, 3, 0, 153, 7, 0, 0, 4, 3, 17, 0, 0, 3, 10, 0, 1, 10, 46, 0, 0, 12, 14, 3, 0, 3, 1, 2, 0, 469, 27, 0, 0, 2], [0, 44, 302, 0, 1, 349, 311, 26, 0, 5, 0, 2, 0, 290, 4, 1, 0, 5, 3, 141, 0, 0, 1, 3, 0, 5, 17, 60, 0, 0, 6, 5, 8, 0, 0, 0, 5, 0, 16, 44, 0, 2, 17], [0, 67, 89, 0, 0, 282, 87, 24, 0, 4, 0, 3, 0, 305, 3, 2, 2, 7, 1, 583, 0, 0, 0, 4, 0, 7, 15, 58, 0, 0, 5, 13, 4, 0, 12, 0, 4, 1, 109, 38, 1, 0, 6], [0, 137, 73, 0, 0, 17, 71, 26, 0, 4, 0, 5, 0, 353, 3, 0, 2, 12, 1, 122, 0, 0, 0, 5, 0, 2, 35, 39, 0, 0, 4, 4, 18, 0, 7, 0, 7, 0, 44, 14, 1, 1, 10], [0, 1389, 207, 0, 1, 0, 216, 12, 0, 0, 0, 2, 0, 483, 2, 1, 1, 17, 1, 23, 0, 0, 2, 4, 0, 0, 38, 51, 0, 0, 5, 12, 656, 0, 2, 0, 5, 0, 5, 13, 0, 0, 7], [0, 547, 187, 0, 0, 1, 192, 19, 0, 1, 0, 1, 0, 208, 1, 0, 0, 45, 1, 14, 0, 0, 0, 4, 0, 2, 43, 36, 0, 0, 2, 4, 180, 0, 0, 0, 4, 0, 2, 8, 0, 0, 6], [0, 78, 202, 0, 1, 6, 210, 13, 0, 0, 0, 1, 0, 23, 4,
[…]

### 2015-01-13 paper.tex
updated paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\documentclass[twoside]{article}

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

\newtheorem{claim}{Claim}

\usepackage{graphicx}
\graphicspath{ {figures/} }

\usepackage[usenames,dvipsnames]{color} % Required for specifying custom colors and referring to colors by name

\definecolor{MyRed}{rgb}{0.6, 0.0, 0.0} 
\definecolor{MyGreen}{rgb}{0.0,0.4,0.0} % Comment color
\definecolor{MyBlue}{rgb}{0.0, 0.0, 0.6}

%\setlength\parindent{24pt}

\usepackage[pdftex]{hyperref} % For hyperlinks in the PDF
\hypersetup{
  colorlinks=true,
  linkcolor=MyBlue, 
  citecolor=MyRed,
  urlcolor= MyGreen
}

% make hyperlinks bold
\newcommand{\aref}[1]
 {\textbf{\autoref{#1}}}

\newcommand{\nref}[1]
 {\textbf{\nameref{#1}}}

\newcommand{\cc}[1]
 {\textbf{\cite{#1}}}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Estimating Trending Topics on Twitter with Small Subsets of the Total Data}} % Article title

\author{
\large
\textsc{Evan Miller}\\[2mm] 
\textsc{Kiran Vodrahalli}\\[2mm]
\textsc{Albert Lee}\\[2mm]
\vspace{-5mm}
}
\date{January 13, 2015}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction} \label{sec:Intro}

% introduction 
Since its founding in $2006$, Twitter has grown into a corporation dominating a significant
proporition of social media today. In recent years, it has become interesting to analyze tweets
to determine interesting phenomena ranging from the viral spread of news to the detection of earthquakes \cc{Burks:2014}. The volume of existing tweets is prohibitively large for standard methods of analysis, and requires approaches that are able to either store large amounts of data in memory for fast access (including parallel approaches to data processing) or approaches that only require sublinear memory while retaining accuracy guarantees. In this paper, we investigate a problem utilizing the latter method.

\subsection{Problem Statement}

% give clear definition of problem and motivation here
We would like to create an online algorithm that provides a real-time estimate of the frequencies of hashtags on Twitter time series data.
Since we want to be able to run this algorithm in real time, we would like for it to be space-efficient: that is, it should not be required to store a large number of hashtag counts at any given time. Ideally, the algorithm's space complexity should be sublinear in the number of hashtags seen. 
Therefore, our estimate should not require knowledge of the exact frequency history of hashtags in tweets.

More specifically, we will attempt to approximate the $k$ most popular Twitter hashtags in the time intervals of length $y$ in order to tell what is trending during that time period.
The idea is that estimating the top $k$ hashtags in some time interval provides a model for the topics that are trending on Twitter in that time period.

\subsection{Previous Work} \label{sec:PreviousWork}

% describe past work in the areas here
Estimating frequencies with high probability and low space complexity has been an interest in
algorithms development for a fair amount of time. The Count-Min sketch, developed in the early $2000$s,
provided an approximate data structure able to maintain counts of keys seen in an online setting in space sublinear to the number of keys, with the size of the structure instead dependent on parameters tunable for different accuracy and guarantee settings \cc{Cormode:2005}.

More recently in $2012$, Matusevych et al. developed the Hokusai structure, which gives frequency counts for keys over a range of time \cc{Matusevych:2012}. The rough idea behind this approach is that in the distant past, we should only care about heavy-hitters, i.e. hashtags with high frequencies in order to estimate the likelihood that the hashtag is trending again. The goal of the time-aggregated Hokusai system is to store older data at decreased precision since the older data also has decreased value in the computation.  The time-aggregated Hokusai system works by storing aggregate data in Count-Min sketches each with a $2^i$ day resolution for the past $2^i$ days.  Each of these Count-Min sketches computes $d$ hash functions from $\{0,1\}^*$ to $\{0, 1, ..., m-1\}$.

Analyzing what is trending on Twitter is a more recent research area. Twitter itself must give its users hashtags which it deems trending, to both spread the hashtag and to provide an awareness of what is going on in the world to its users. Knowledge of what is trending and how that distribution looks over time is also useful for the business model, which relies on ads to a certain extent.
Twitter itself stores all of its frequency data, and looks at recent spikes while conditioning on all of the past frequency data in order to determine estimates of trending likelihood.
In $2012$, MIT Professor Devavrat Shah built a machine-learning algorithm to predict which topics will trend on Twitter an hour and a half before they do, with $95\%$ accuracy, though the datasets they tested their algorithm on were small\cc{MITNews2012}.

As a preliminary starting point for our approach to approximating the past frequencies, we will utilize concepts from the Hokusai paper to generalize the Count-Min Sketch scheme to time-series data \cc{Matusevych:2012}.

\subsection{Our Approaches}

Our approach was to explore whether or not adding historical data would improve the accuracy of 
trending hashtag detection. To that end, we implemented two algorithms: one called the naive algorithm, the other called the history-sensitive algorithm. We wanted to test whether an algorithm that uses
the history (in condensed form) to help understand what is trending at the current time could outperform an algorithm with a very short term memory.

The reason why the history is important is as follows: Some hashtags are constantly present. For example, in our experiments, the hashtags $\#$porn, $\#$gameinsight, $\#$teamfollowback had nearly always had a high frequency in the current time unit. This fact is not surprising given the nature of the first hashtag. The second hashtag, $\#gameinsight$, is the name of popular video game maker that people tweet when they win achievements in the game. Finally, $\#teamfollowback$ is a method Twitter users apply to gather Twitter followers quickly. The rule is if you follow that tag, you follow people who post that tag. Since these sorts of tags are consistently present, they should not be consider trending: they are not new. Therefore, we need to filter these out with the history.
The necessity of maintaining approximate counts for a longer history depends on exactly how constant these consistently-appearing tweets actually are.

We had two different algorithms for finding out which hashtags were trending in real time, both space efficient.
The naive algorithm does not take into account the history and simply looks at a $y$-unit before the current $y$-unit interval to find the current trending hashtags. To determine whether a hashtag was trending, we tested

\begin{equation}
\label{eq:threshold}
\frac{\textbf{freq}(\textnormal{hashtag}, \textnormal{present } y \textnormal{-unit})}{\textbf{freq}(\textnormal{hashtag}, \textnormal{past } y \textnormal{-unit})} > c
\end{equation}

where $c$ is a threshold we picked to determine whether the hashtag was considered trending. We believe that in order to accurately determine trending hashtags, we will need to use more of the past in order to eliminate the hashtags which are ever-present, and therefore not trending.

The second approach differs in that we do not store all the data of the past, but instead use several Count-Min Sketches to approximate the past frequencies. We call this approach the history-sensitive algorithm.
The idea is to remember the past with less accuracy to save space, while weighting the frequency counts of a hashtag in the past appropriately. We used the rule of thumb that the more often a hashtag shows up in the past, the less likely it is trending. However, we also discount the importance of the past by weighting the times long ago less and less as time moves forward.

%------------------------------------------------

\section{Describing the History-Sensitive Algorithm and Data Structure}

% how did we approach it 
We separate the problem of finding trending topics on Twitter into two parts.  First, we need to maintain a data structure that efficiently stores data about all occurrences of every hashtag seen in the past.  We also maintain a separate data structure that allows us to quickly gather information about the most recent hashtags seen.

We want the former data structure to be very space efficient since it must store data about a very large dataset.  For this structure, space efficiency is more important than accuracy since small deviations in such a large dataset should not be significant because the deviations in past data should not greatly affect what is currently trending.

For the latter data structure, accuracy is more important than space efficiency since the structure contains data which more closely relates to which topics are currently trending and the size of the dataset is much smaller.

\subsection{Data Structures}
% -- we maintain datastructure in two parts: History, and Present
\subsubsection{History Data Structure}

To store data about all occurrences of every hashtag seen in the past, we use a modified version of the time-aggregated Hokusai system \cc{Matusevych:2012}, which is an extension of the Count-Min sketch. We previously described this data structure in \nref{sec:PreviousWork}. 
To the Hokusai structure we add another Count-Min sketch that combines the information from the Hokusai Count-Min sketches. We call this external Count-Min sketch the Kernel, since it acts as a weighting function on the CM sketches in the Hokusai structure. Its role is to depreciate the value of older data.
Denoting the CM sketches of the Hokusai structure as a vector $\textbf{M} = \{\overline{M}, M^0, M^1, ..., M^{{\lfloor {\log(T)} \rfloor}}\}$, where $\overline{M}$ is the Count-Min sketch storing the data of a unit time step (of size $z$), $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ time resolution, and $T$ is an upper bound on the number of $z$-units stored by the data structure. Then, the kernel function is given by

\begin{equation} 
\label{eq:kernel}
\textit{k}(\textbf{M}) = \overline{M} + \sum\limits_{j=0}^{{\lfloor \log⁡(T) \rfloor}} \frac{M^j}{2^j}
\end{equation}

For each hashtag, the kernel sketch stores a positive value between $0$ and $2$ (though typically $\leq 1$) that approximates how often the hashtag showed up in the past. In \nref{sec:Correctness} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i > 0$ days ago with weight approximately $\frac{1}{i}$.

We will refer to the combined data structure of Hokusai and the Kernel as the History.

\subsubsection{Current-Window Data Structure}

Our data structure for holding hashtags seen in the last $y$-length time period consists of three components: a max Fibonacci heap, a queue, and a hash table.  We refer to these components collectively as the Current-Window.
The keys for the hash table are the hashtags, and the values stored in the table are (frequency, pointer to corresponding node in heap) pairs.
The queue contains (hashtag, timestamp) pairs, each of which is inserted upon seeing a hashtag in the input stream.
Then, we need to define a heap priority function, which defines what trending means for the whole data structure. 
For a specific hashtag, we define the priority function to be

\begin{equation} 
\label{eq:priority}
\textbf{priority}(\textnormal{hashtag}) = \frac{\textbf{freq}(\textnormal{hashtag}, \textnormal{present } y \textnormal{-unit})}{\textbf{History}[\textnormal{hashtag}]}
\end{equation}

The idea is that again, we want to be inversely proportional to the history, here corrected by more knowledge and clever approximation schemes to save space.

\subsection{Algorithm Pseudocode}

\subsubsection{Updating History Data Structure}

Algorithm \ref{alg1} describes the necessary steps to maintain the History data structure as new input is provided. We perform time aggregation to fill the Kernel, $K$.

\begin{algorithm} 
\caption{Update History} \label{alg1}
\begin{algorithmic}[1]
\ForAll {$i$}
	\State Initialize Count-Min sketch $M^i = 0$
\EndFor
\State Initialize $t = 0$
\State Initialize Count-Min sketches $\overline{M} = 0$ and $K = 0$
\While {data arrives}
	\State Aggregate data into sketch $\overline{M}$ for current $z$-unit while also adding this data to $K$
	\State $t \leftarrow t + 1$ (increment counter)
	\State $K \leftarrow K - \overline{M}$
	\For {$j = 0$ to argmax \{$l$ where $t$ mod $2^l = 0$\}}
		\State $K \leftarrow K + 2^{-j}(\overline{M} - M^j)$
		\State $T \leftarrow \overline{M}$ (back up temporary storage)
		\State $\overline{M} \leftarrow \overline{M} + M^j$ (increment cumulative sum)
		\State $M^j \leftarrow T$ (new value for $M^j$)
	\EndFor
	\State $\overline{M} \leftarrow 0$ (reset aggregator)
\EndWhile
\end{algorithmic}
\end{algorithm}

\subsubsection{Updating heap and hash tables}

This is 2.

\begin{algorithm} 
\caption{Update Current-Window} \label{alg2}
\begin{algorithmic}[1]
\While {data arrives}
	\If {the current $z$-unit is different than that of the last hashtag seen}
		\State Do all the end-of-$z$ aggregation for the History structure as detailed in 
		Algorithm \aref{alg1}.
		\ForAll {elements in the Fibonacci heap}
			\State look up the hashtag corresponding to this node in the hash table
			\State Update the key of the node to: \[\frac{\textnormal{frequency in last } y-\textnormal{length time period found at the table entry}}{\textnormal{new value in History data structure}}\]
		\EndFor
	\EndIf
	\If {queue is not empty}
		\State Peek at end of queue.
		\If {the timestamp + $y$ is before the current time}
			\State Look up the hashtag in the hash table and decrement the stored frequency.
			\If {the frequency is now 0}
				\State Delete the node in the heap pointed to by this entry in the table.
				\State Delete this entry in the hash table.
			\Else	
				\State Update the key of this node pointed to by this entry in the table to the proper value given the new frequency.
			\EndIf
		\EndIf
	\EndIf
	\If {hashtag is in hash table}
		\State Increment the frequency stored at that entry.
		\State Update the key of the node in the Fibonacci heap.
	\Else
		\State Insert a new node into the Fibonacci heap with the appropriate key and value. 
		\State Insert the hashtag into the hash table with a pointer to this node.
 	\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

\subsubsection{Finding the trending hashtags}

This is 3.

\begin{algorithm} 
\caption{Top $k$ trending hashtags} \label{alg3}
\begin{algorithmic}[1]
\State Perform $k$ delete-max operations on the Fibonacci heap, storing each of the nodes deleted in a list $L$.
\ForAll {nodes $N$ in $L$}
	\State Announce that the hashtag associated with $N$ is trending.
	\State Insert $N$ into the Fibonacci heap.
\EndFor
\end{algorithmic}
\end{algorithm}

%------------------------------------------------

\section{Analyzing the History-Sensitive Algorithm}

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Symbol & Meaning                                                           \\ \midrule
d      & Number of hash tables for each Count-Min sketch                        \\
m      & Size of each hash table in each Count-Min sketch                  \\
s      & Number of distinct hashtags in the Current-Window                 \\
z      & Time resolution of the unit-sized Count-Min sketches                   \\
T      & Upper bound on the number of $z$-units of data stored in History                   \\
x      & Total number of hashtags in the Current-Window                   \\
y      & Time interval of data contained in the Current-Window                   \\ \bottomrule
\end{tabular}
\caption{Variables referenced in this section}
\end{table}

\subsection{Correctness} \label{sec:Correctness}

This algorithm finds the $k$ hashtags that have the maximum value of (frequency in last $y$-unit) / (value in History data structure).

\begin{claim}
The value for hashtag $x$ in the aggregate Count-Min sketch of the History data structure is within a factor of 4 of $\overline{M}(x) + \sum\limits_{i=1}^T \frac{1}{i}*$ (value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring i days ago).
\end{claim}
\begin{proof}
First, we use Theorem $4$ in \cc{Matusevych:2012} which states that ``At $t$, the sketch $M^j$ contains statistics for the period $[t - \delta, t - \delta - 2^j]$ where $\delta = t \mod 2^j$.''

Let $b$ be the location in the aggregate Count-Min sketch containing the value returned when $x$ is queried.

Let $h$ be any instance of any hashtag that appeared on day $p$ such that seeing $h$ incremented counters in position $b$.

Case 1: $p=t$

Then the statistics for $h$ are recorded in $\overline{M}$ and are not in any $M^j$.

Case 2: $2^i > t-p \geq 2^{i-1}$ for some $i > 0$

By Theorem 4, for all $j \leq i - 2$, $M^j$ does not contain statistics for $p$ since $p \leq t - 2^{i-1} \leq t - \delta - 2^{i-2}$.

Therefore, the increments that occurred in the Count-Min sketch for hashtags occurring $i$ days ago contribute at most $\sum\limits_{j = i - 1}^{T} 2^{-j} < 2^{2-i}$ to the value in position $b$.

Let $q$ be the largest $j$ such that $t - \delta - 2^j \geq p$

Then $p \leq t - \delta - 2^q$.  Let $\lambda = t \mod 2^{q+1}$.  Then $\lambda = \delta$ or $\lambda = \delta + 2^q$.

Since $q$ is the largest $j$ such that $t - \delta - 2^j \geq p$, $p > t - \lambda - 2^{q+1}$.

Also, $p \leq t - \delta - 2^q \leq t - \lambda$, so $M^{q+1}$ contains statistics about $h$.

For all $j \geq i$, $t - \delta - 2^j < t - 2^j < p$, so $q \leq i - 1$.

Thus, incrementing the counter in $M^{q + 1}$ contributed at least $2^{-i}$ to the value in position $b$.

Thus, the contributions to the sum are within a factor of 4 of $\frac{1}{t-p}$.

Therefore, summing over all hashtags that increment counters in position $b$ gives $\overline{M}(x)$ for all hashtags that occurred on day $t$, and within a factor of 4 of $\sum\limits_{i=1}^T \frac{1}{i}*$(value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring $i > 0$ days ago). 
\end{proof}
This value is approximately (frequency in last $y$-unit) / (freq today + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag $i$ $z$-units ago)).
This seems to be a desirable function to maximize since it finds hashtags that are common in the last $y$-unit that have been comparatively infrequent in the past.  This function is good since it especially emphasizes hashtags that are different than those seen in the past few days.  This ensures that the same items do not stay trending for too long.

\subsection{Runtime Analysis}

Processing the insertion of a hashtag takes the following time.  It takes amortized O(d) time to update the History.  It takes expected O(1) time to check if in hash table.

If it is, it requires O(1) time to increment the frequency, O(d) time to compute the new key, and O(1) amortized time to update the key since it will be nondecreasing.

Otherwise, it requires O(d) time to compute the key value, O(1) time to insert the new node in the heap, and O(1) time to insert into the hash table.

Thus, our algorithm takes O(d) amortized time + expected O(1) time to process a new hashtag.

\begin{table}[h]
\centering
\label{my-label}
\begin{tabular}{@{}ll@{}}
\toprule
Case                                                                                                   & Amortized Time          \\ \midrule
Processing the insertion of a new hashtag                                                              & $O(d)$                  \\
\begin{tabular}[c]{@{}l@{}}Processing the removal of a hashtag from \\ the Current-Window\end{tabular} & $O(d + \log(s))$        \\
\begin{tabular}[c]{@{}l@{}}Updating History and Current-Window \\ at the end of a $z$-unit\end{tabular}     & $O(md + ds + s\log(s))$ \\
Querying for the top $k$ trending items                                                                & $O(k\log(s))$           \\ \bottomrule
\end{tabular}
\caption{Time analysis summary}
\end{table}

Processing the removal of a hashtag from the Current-Window takes the following time.  It takes O(1) time to verify that the queue is not empty.  It takes O(1) time to look at the end of the queue and verify that the timestamp + $y$ is before the current time.  It takes O(1) time in expectation to look up this hashtag and decrement its frequency.  Then, it takes O(1) time to check if the frequency is 0.

If so, it takes O(log(s)) amortized time to delete the node in the heap and O(1) time to delete the entry in the hash table.

Otherwise, it takes O(d) amortized time to compute the new key for the hash table and O(log(s)) amortized time to update the heap given this key.

Thus, our algorithm requires O(log(s)) amortized time + expected O(1) time  + O(d) time to remove a hashtag from the Current-Window.

The end-of-day updates to the History and the resulting updates to the heap take the following time.  By Lemma 5 of the Hokusai paper, the amortized time required to do all end-of-day aggregation is O(md).  Then, for each of the s nodes in the heap, it takes O(d) time to compute each updated key and O(log(s)) amortized time to update the heap given the new key.

Thus, it takes O(md + slog(s)) amortized time + O(ds) time to do all necessary updates at the end of the day.

Querying the data structure for the top k trending items takes O(k*log(s)) amortized time for delete-max operations, O(k) time to announce that these items are trending, and O(k) amortized time to reinsert these nodes into the heap.

Thus, it takes O(k*log(s)) amortized time to determine what’s trending.

\subsection{Spatial Analysis}

The History requires O(md) space for each Count-Min sketch, so it requires a total of O(md*log(T)) space.

Each node in the heap requires a constant amount of space, so the heap requires O(s) space.

The hash table always contains at most s entries with each entry requiring a constant amount of space.  Also, in order to maintain an expected O(1) lookup time, the hash table needs to have O(s) bins.  Thus, the hash table requires O(s) space.

The queue requires an entry for every hashtag still in the current window, so it requires O(x) space.

Thus, everything requires O(md*log(T) + x) space since s $<$ x.

%------------------------------------------------

\section{Design Choices}

% y is 3 hours, z is 1 day
\subsection{Choosing the parameters $y$ and $z$}
For $y$, the unit time period we attempted to find the trending hashtags for, we chose a $3$-hour interval.
For $z$, the unit time period for the History data structure, we chose day-length intervals. These values made sense because the testing data we used occurred over a $30$-day interval, and we wanted to avoid too fine a temporal resolution so that we could run our code on the data in reasonable time.

\subsection{Parameters of the History Data Structure}
		pick some parameters here
		n = 6, m = 3500 (size of hash tables), d = 20 (number of tables)
		n: we do it for a month, so $\sim 2^{(6-1)}$ days is definitely enough
		m = ceiling(e/epsilon), d = ceiling(ln(1/delta))
		error in query is in factor epilson with probability 1 - delta
		for d = 20, we have with high probability (1 - 1/$e^{20}$)
		we have error factor of e/3500 of the total number of hashtags
		total number of hashtags is $\sim$ 15 million
		thus error bounds are in range $\sim$ 12000 on a given count
		this error provides a 'good' amount of smoothing

\subsection{Choosing the Flavor of Heap}

We chose a Fibonacci heap in the hopes that there would be a larger number of $\textbf{decrease-key}$ operations, which has a better theoretical bound in Fibonacci heaps.

%------------------------------------------------

\section{Testing Procedure}

We downloaded the \href{https://archive.org/details/twitterstream}{September $2014$ archive of Twitter data } to serve as test data for our algorithms \cc{Twitter2014}. We then parsed the data into (timestamp, hashtag) format
for all thirty days worth of data. Using these files as pseudo-real-time input, we used the last week of the data as a benchmark for comparison between the naive and the history-sensitive algorithms. We divided the data into three-hour chunks, and queried for the top ten hashtags in each of the three-hour time blocks
in the last week of the data using both algorithms. Note that the history-senstive algorithm was required to parse through the first three weeks as well in order to build its history, whereas the naive algorithm was only required to pay attention to six-hour chunks at a time, and thus never needed to look a bit beyond the scope of that week.

We used the Jaccard similarity to arrive at a similarity score between the two methods. We also investigated the hashtags that both algorithms came up with for a given $3$-hour interval, and verified
that they were in fact trending at that time using news and Twitter media.

%------------------------------------------------

\section{Results}

\subsection{Performance Measurements}

We use the space complexities of the data structures and the sizes of the 
respective inputs to determine approximately how much space is used by each
approach, and compare them.

\subsubsection{Naive Algorithm}

Space measurements (running on one week's worth of hashtags) (estimate 6 hours worth of hashtags)

Time (running over one week): $\sim 15$ minutes

\subsubsection{History-Sensitive Algorithm}

(we had to build a history over the first three weeks of the month
 in order to determine the History data structure's values for the week we ran
 the algorithm over)

Space measurements (running on one week's worth of hashtags):

Time (running over one week): $2 - 3$ hours

\subsection{Are the Algorithm Outputs Actually Trending?}

\subsubsection{The Intersection Distribution}

First, we see how many hashtags were marked as trending 
by both the naive and history-sensitive algorithms in the same time slots.
For each $3$-hour interval, we calculate the Jaccard Similarity, a value
in the range $[0, 1]$, given by

\begin{equation}
\label{eq:jaccard}
J(A, B) = \frac{|A \cap B|}{|A \cup B|}
\end{equation}

\noindent where $A$ and $B$ are the sets of top $10$ hashtags for a given $3$-hour interval
for each algorithm. 
We then calculate the mean and standard deviation of the Jaccard similarity
over the $56$ different $3$-hour intervals of the last week of September $2014$.

This value gives us a metric for understanding how similarly the two algorithms performed.

We plot it below in the histogram shown in \aref{fig:jaccard_sims}. The $x$-axis is the Jaccard similarity, and the $y$-axis is the count. The mean was $0.47$, and the standard Deviation was $0.18$. From the plot, we can see that the distribution of the Jaccard similarities is roughly a centered normal distribution around $0.5$. A Jaccard similarity of $0.5$ means about half the hashtags are the same in each $3$-hour time interval between the two algorithms.

\begin{figure}
\centering
\includegraphics[scale = .4]{jaccard_sims}
\caption{Distribution of Jaccard Similarities}
\label{fig:jaccard_sims}
\end{figure}

\subsubsection{Checking the Algorithm-Defined Trends Against Real Life}

In \aref{fig:news_vis}, we have superimposed pictures from 
news articles that the tweets describe on a time axis, as
the hashtag referencing the article starts trending. The Jaccard similarities 
for each of these three-hour intervals were all above $0.5$, so both algorithms
found the hashtags that we graph in the figure.
We checked that the news was released (via Yahoo! News, Twitter, Google News, ESPN, and others)
at around the time our algorithms say the hashtag relating to the news
begins to trend. A fair number of the trending hashtags related to sports and TV shows.

As we can see in the figure, our algorithm detected the start of the TV show "How to Get Away With Murder", which was released on September $25$, $2014$. Another TV-related hashtag was $\#dailyshowgonetoofar$, when Jon Stewart called One Direction member Zayn Malik a terrorist (on the $26^{th}$) and caused a huge outrage. Sports events that occur real-time on TV also have the trending profile. There were three in this section of six three-hour intervals. First, Derek Jeter hit a walk-off homerun in the final time at bat on the Yankees, and naturally on Twitter, fans were excited. This event occurred on the $25^{th}$. The Ryder Cup is a major golf tournament that began on the $26^{th}$ and lasted through the $28^{th}$, and its beginning launched a wave of tweets. The last sports-related hashtag was the beginning of the ToyotaEMR car race that took place in India on September $26^{th}$. Finally, $\#blowthewhistle$ related to a kids cancer campaign that had lasted throughout September, which perhaps got a boost as September ended to raise some last minute funds.

\begin{figure} 
\centering
\includegraphics[scale = .4]{Timeline}
\caption{News Visualization of Trending Tweets}
\label{fig:news_vis}
\end{figure}

%------------------------------------------------

\section{Discussion}

\subsection{Comparison of Naive and History-Sensitive Algorithms}

observed that history-sensitive 
did about the same in terms of getting trending tweets (based on Jaccard), 
but naive uses much less space.

\subsection{Applicability to Real-Time Data}

do well to get month's worth of data in a few hours (2-3)

implementation wouldnt run in python, do it in C or something

for real data from firehose (all of it), we would use 
better infrastructure dedicated to processing all the data
would use parallel setup (this scheme is adaptable to parallel
setup)

seems reasonable if we have proper data storage setup.

%------------------------------------------------
\section{Future Work} \label{sec:Future Work}

With respect to the algorithms specifically, in the future we would want to adapt our approach to the history-sensitive algorithm and play with the parameters a bit more. We would also like to try different kernel functions to see how they performed. Another approach to utilizing the history information would be instead of weighting by inverse exponentially weighted frequency in the past for a given hashtag, attempt to calculate some score of periodicity for the hashtag. If the hashtag has low periodicity over a given time scale, then even if it appears multiple times in the past (because the past is long), our algorithm will avoid discarding it as non-trending.

Then, we would like to try adapating our algorithms to real-time data: that is, hook up our algorithms to a significant portion of the Twitter firehose, after setting up the necessary hardware to process that amount of data. In order to aid with the processing, we would also ideally use a distributed system and parallelize
our code for better performance.

We also know that an estimation the top $k$ hashtags in a given time interval could provide a topic model for tweets. It could be interesting to use the top $k$ hashtags to build a time-sensitive topic model and analyze quantities including the speed of topic change, the variety of topics, and so on.

Even though not perhaps irrelevant to the specific task of identifying trending hashtags, 
the approach of the history-sensitive algorithm may be useful in other domains.
Future work could also investigate the performance of the history-sensitive algorithm on other data streams of interest, for instance, Wikipedia edits -- our goal would be to estimate which Wikipedia topics are being edited the most at any given time interval of some length. Another interesting application of this algorithm to test would be the real-time estimatation of the hottest selling stocks on Wall Street to see if either of these settings results in a significant improvement over the naive algorithm.

\section{Appendix I: Code and Visualizations} \label{sec:Appendix_code_viz}

We provide a link to \href{https://github.com/kiranvodrahalli/cos521/}{all our code}, as well as to an online hosting of the \href{http://www.princeton.edu/~awlee/trends.html}{frequency visualization}.

In our code, we made use of two libraries written by other people, which we modified. One of these was a Count-Min sketch implementation in Python, and another was a Fibonacci heap implementation, also in Python. These program files are in our codebase, with appropriate citations at the top of each document. Furthermore, we used multiple standard Python libraries in our code.

\section{Appendix II: Top-$k$ Hashtags in $6$ Interesting Time Chunks} \label{sec:Appendix_topk}

To see all $56$ $3$-hour intervals for both algorithms, see the results folder of our \href{https://github.com/kiranvodrahalli/cos521/}{Github}. We display the top $10$ hashtags for each $3$-hour interval for both the naive algorithm and the history-sensitive algorithm, along with their heap priorities. Note that the heap priorities span a reasonable range from $0$ to $1$, and are not clustered around $1$ or $0$, implying the algorithms are actually differentiating the hashtags as trending and not trending.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Burks2014]{Burks:2014}
Burks, L., Miller, M., and Zadeh, R. (2014). 
\newblock RAPID ESTIMATE OF GROUND SHAKING INTENSITY BY COMBINING SIMPLE EARTHQUAKE CHARACTERISTICS WITH TWEETS. 
\newblock {\em Tenth U.S. National Conference on Earthquake Engineering Frontiers of Earthquake Engineering}. Retrieved from \href{http://www.stanford.edu/~rezab/papers/eqtweets.pdf}{http://www.stanford.edu/~rezab/papers/eqtweets.pdf}

\bibitem[Cormode2005]{Cormode:2005}
Cormode, G., and Muthukrishnan, S. (2005). 
\newblock An improved data stream summary: The count-min sketch and its applications. 
\newblock {\em Journal of Algorithms, 55}(1), pp. 58-75.

\bibitem[MITNews2012]{MITNews2012}
Hardesty, Larry. 
\newblock {\em "Predicting What Topics Will Trend on Twitter."} 
\newblock MIT News Office. MIT, 1 Nov. 2012. Web. 13 Jan. 2015. \href{http://newsoffice.mit.edu/2012/predicting-twitter-trending-topics-1101}{http://newsoffice.mit.edu/2012/predicting-twitter-trending-topics-1101}.

\bibitem[Matusevych2012]{Matusevych:2012}
Matusevych, S., Smola, A., and Ahmed, A. (2012).
\newblock Hokusai--Sketching Streams in Real Time.
\newblock {\em UAI '12: 28th conference on Uncertainty in Artificial Intelligence}, pp. 594-603.

\bibitem[Twitter2014]{Twitter2014}
Twitter, Inc. (2014). 
\newblock {\em Archive Team JSON Download of Twitter Stream 2014-09}. 
\newblock Retrieved from \href{https://archive.org/details/twitterstream}{https://archive.org/details/twitterstream}.

\bibitem[Image1]{Image1}
\newblock BlowTheWhistle. Retrieved from \href{https://safesite.4agoodcause.com/curing-kids-cancer/donate/template_files/Blow_The_Whistle.jpg}{this link.}

\bibitem[Image2]{Image2}
\newblock How to Get Away with Murder.  Retrieved from \href{http://static.east.abc.go.com/service/image/index/id/65068ab1-d543-4958-a2cc-ee366bd807db/dim/600x800.jpg}{this link.}

\bibitem[Image3]{Image3}
\newblock One Direction terrorsist joke goes south for Daily Show.  Retrieved from \href{https://pbs.twimg.com/media/B52-mq-IQAEIqxz.jpg}{this link.}

\bibitem[Image4]{Image4}
\newblock Adrenna pic. Retrieved from \href{http://indiainf1.com/wp-content/uploads/2014/09/MRF-FF1600-BIC-28SEp2014-Adrenna-pic-300x184.jpg}{this link.}

\bibitem[Image5]{Image5}
\newblock Rory Mac.  Retrieved from \href{http://i.telegraph.co.uk/multimedia/archive/03050/rorymac_3050367b.jpg}{this link.}

\bibitem[Image6]{Image6}
\newblock Derek Jeter closes his Yankee Stadium finale with walk-off single.  Retrieved from \href{http://www.foxsports.com/mlb/story/derek-jeter-closes-yankee-stadium-finale-with-walk-off-single-new-york-yankees-beat-baltimore-orioles-092514}{this link.}

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2015-01-13 paper.tex
update paper

@@
 
-Our data structure for holding hashtags seen in the last $y$-length time period consists of three components: a max Fibonacci heap, a queue, and a hash table.
+Our data structure for holding hashtags seen in the last $y$-unit consists of three components: a max Fibonacci heap, a queue, and a hash table.
 We refer to these components collectively as the Current-Window.
 The keys for the hash table are the hashtags, and the values stored in the table are (frequency, pointer to corresponding node in heap) pairs.
+The hash table exists so that we can keep track of exact frequency counts for each $y$-unit.
 The queue contains (hashtag, timestamp) pairs, each of which is inserted upon seeing a hashtag in the input stream.
-Then, we need to define a heap priority function, which defines what trending means for the whole data structure. 
+The purpose of the queue is to determine when a hashtag count is too old to be maintained in the present $y$-unit window.
+Finally, we have a Fibonacci heap to prioritize the hashtags based on how trending they currently are.
+We provide a heap priority function, which defines what trending means for the whole data structure. 
 For a specific hashtag, we define the priority function to be
@@
 
-Algorithm \ref{alg1} describes the necessary steps to maintain the History data structure as new input is provided. We perform time aggregation to fill the Kernel, $K$.
+Algorithm \aref{alg1} describes the necessary steps to maintain the History data structure as new input is provided.
+We perform time aggregation to fill the Kernel, $K$.
+This algorithm is only called when a $z$-unit has passed since the last time the algorithm was called.
 
@@
 
-This is 2.
+We run Algorithm \aref{alg2} in the main loop, every time a single piece of data arrives.
+For each new piece of data, we also check whether the data we are storing has grown 
+outdated -- that is, older than a $y$-unit before the most current piece of data arrived.
 
@@
 
-This is 3.
+We use Algorithm \aref{alg3} to calculate the top $k$ hashtags by querying the heap.
+The heap is the structure responsible for maintaining the priority of the various hashtags, and provides an efficient way of implementing a changing priority queue.
 

### 2015-01-13 paper.tex
added smaller font to fit on 11 pages
added all logic to paper
almost final draft. need to clean up the algorithms section 2 and section 3.1

@@
 %----------------------------------------------------------------------------------------
+\small{
 
@@
 Symbol & Meaning                                                           \\ \midrule
-d      & Number of hash tables for each Count-Min sketch                        \\
-m      & Size of each hash table in each Count-Min sketch                  \\
-s      & Number of distinct hashtags in the Current-Window                 \\
-z      & Time resolution of the unit-sized Count-Min sketches                   \\
-T      & Upper bound on the number of $z$-units of data stored in History                   \\
-x      & Total number of hashtags in the Current-Window                   \\
-y      & Time interval of data contained in the Current-Window                   \\ \bottomrule
+$d$      & Number of hash tables for each Count-Min sketch                        \\
+$m$      & Size of each hash table in each Count-Min sketch                  \\
+$s$      & Number of distinct hashtags in the Current-Window                 \\
+$z$      & Time resolution of the unit-sized Count-Min sketches                   \\
+$T$      & Upper bound on the number of $z$-units of data stored in History                   \\
+$x$      & Total number of hashtags in the Current-Window                   \\
+$y$      & Time interval of data contained in the Current-Window                   \\ \bottomrule
 \end{tabular}
@@
 
+\newpage
 \subsection{Correctness} \label{sec:Correctness}
 
-This algorithm finds the $k$ hashtags that have the maximum value of (frequency in last $y$-unit) / (value in History data structure).
+Our history-sensitive algorithm finds the $k$ hashtags that have the maximum value of 
+$
+\frac{\textbf{freq}(\textnormal{hashtag}, \textnormal{present } y \textnormal{-unit})}{\textbf{History}[\textnormal{hashtag}]}
+$.
 
 \begin{claim}
-The value for hashtag $x$ in the aggregate Count-Min sketch of the History data structure is within a factor of 4 of $\overline{M}(x) + \sum\limits_{i=1}^T \frac{1}{i}*$ (value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring i days ago).
+The value for hashtag $x$ in the aggregate Count-Min sketch of the History data structure is within a factor $4$ of $\overline{M}(x) + \sum\limits_{i=1}^T \frac{1}{i}*$ (value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring $i$ $z$-units ago).
 \end{claim}
@@
 Let $b$ be the location in the aggregate Count-Min sketch containing the value returned when $x$ is queried.
+Let $h$ be any instance of any hashtag that appeared on $z$-unit $p$ such that seeing $h$ incremented counters in position $b$.
 
-Let $h$ be any instance of any hashtag that appeared on day $p$ such that seeing $h$ incremented counters in position $b$.
+\bigskip
 
-Case 1: $p=t$
+\noindent $\textbf{Case 1}$: $p=t$
 
-Then the statistics for $h$ are recorded in $\overline{M}$ and are not in any $M^j$.
+Then, the statistics for $h$ are recorded in $\overline{M}$ and are not in any $M^j$.
 
-Case 2: $2^i > t-p \geq 2^{i-1}$ for some $i > 0$
+\noindent $\textbf{Case 2}$: $2^i > t-p \geq 2^{i-1}$ for some $i > 0$
 
@@
 
-Therefore, the increments that occurred in the Count-Min sketch for hashtags occurring $i$ days ago contribute at most $\sum\limits_{j = i - 1}^{T} 2^{-j} < 2^{2-i}$ to the value in position $b$.
+\bigskip
 
-Let $q$ be the largest $j$ such that $t - \delta - 2^j \geq p$
+Therefore, the increments that occurred in the Count-Min sketch for hashtags occurring $i$ $z$-units ago contribute at most $\sum\limits_{j = i - 1}^{T} 2^{-j} < 2^{2-i}$ to the value in position $b$.
 
+Let $q$ be the largest $j$ such that $t - \delta - 2^j \geq p$.
 Then $p \leq t - \delta - 2^q$.  Let $\lambda = t \mod 2^{q+1}$.  Then $\lambda = \delta$ or $\lambda = \delta + 2^q$.
+Since $q$ is the largest $j$ such that $t - \delta - 2^j \geq p$, $p > t - \lambda - 2^{q+1}$.
+Also, $p \leq t - \delta - 2^q \leq t - \lambda$, so $M^{q+1}$ contains statistics about $h$.
+For all $j \geq i$, $t - \delta - 2^j < t - 2^j < p$, so $q \leq i - 1$. 
+Thus, incrementing the counter in $M^{q + 1}$ contributed at least $2^{-i}$ to the value in position $b$.
+Thus, the contributions to the sum are within a factor $4$ of $\frac{1}{t-p}$.
 
-Since $q$ is the largest $j$ such that $t - \delta - 2^j \geq p$, $p > t - \lambda - 2^{q+1}$.
-
-Also, $p \leq t - \delta - 2^q \leq t - \lambda$, so $M^{q+1}$ contains statistics about $h$.
-
-For all $j \geq i$, $t - \delta - 2^j < t - 2^j < p$, so $q \leq i - 1$.
-
-Thus, incrementing the counter in $M^{q + 1}$ contributed at least $2^{-i}$ to the value in position $b$.
-
-Thus, the contributions to the sum are within a factor of 4 of $\frac{1}{t-p}$.
-
-Therefore, summing over all hashtags that increment counters in position $b$ gives $\overline{M}(x)$ for all hashtags that occurred on day $t$, and within a factor of 4 of $\sum\limits_{i=1}^T \frac{1}{i}*$(value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring $i > 0$ days ago).
+Therefore, summing over all hashtags that increment counters in position $b$ gives $\overline{M}(x)$ for all hashtags that occurred on $z$-unit $t$, and within a factor $4$ of $\sum\limits_{i=1}^T \frac{1}{i}*$(value for $x$ in a Count-Min sketch using the same hash functions for all hashtags occurring $i > 0$ $z$-units ago).
 \end{proof}
-This value is approximately (frequency in last $y$-unit) / (freq today + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag $i$ $z$-units ago)).
+This value is approximately (frequency in last $y$-unit) / (freq current $z$-unit + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag $i$ $z$-units ago)).
 This seems to be a desirable function to maximize since it finds hashtags that are common in the last $y$-unit that have been comparatively infrequent in the past.
-This function is good since it especially emphasizes hashtags that are different than those seen in the past few days.
+This function is good since it especially emphasizes hashtags that are different than those seen in the past few $z$-units.
 This ensures that the same items do not stay trending for too long.
@@
 
-Processing the insertion of a hashtag takes the following time.  It takes amortized O(d) time to update the History.  It takes expected O(1) time to check if in hash table.
+We analyze the time require for processing the insertion of a hashtag. It takes amortized $O(d)$ time to update the History. It takes expected $O(1)$ time to check if the hashtag is in the hash table.
 
-If it is, it requires O(1) time to increment the frequency, O(d) time to compute the new key, and O(1) amortized time to update the key since it will be nondecreasing.
+If it is, it requires $O(1)$ time to increment the frequency, $O(d)$ time to compute the new key, and $O(1)$amortized time to update the key since it will be non-decreasing.
 
-Otherwise, it requires O(d) time to compute the key value, O(1) time to insert the new node in the heap, and O(1) time to insert into the hash table.
+Otherwise, it requires $O(d)$ time to compute the key value, $O(1)$ time to insert the new node in the heap, and $O(1)$ time to insert into the hash table.
 
-Thus, our algorithm takes O(d) amortized time + expected O(1) time to process a new hashtag.
+Thus, our algorithm takes $O(d)$ amortized time $+$ expected $O(1)$ time to process a new hashtag.
 
@@
 
-Processing the removal of a hashtag from the Current-Window takes the following time.
-It takes O(1) time to verify that the queue is not empty.
-It takes O(1) time to look at the end of the queue and verify that the timestamp + $y$ is before the current time.
-It takes O(1) time in expectation to look up this hashtag and decrement its frequency.
-Then, it takes O(1) time to check if the frequency is 0.
+Next we analyze the required time to process the removal of a hashtag from the Current-Window.
+It takes $O(1)$ time to verify that the queue is not empty.
+It takes $O(1)$ time to peek at the end of the queue and verify that the timestamp $ + y$ is before the current time.
+It takes $O(1)$ time in expectation to look up this hashtag and decrement its frequency.
+Then, it takes $O(1)$ time to check if the frequency is $0$.
 
-If so, it takes O(log(s)) amortized time to delete the node in the heap and O(1) time to delete the entry in the hash table.
+If so, it takes $O(\log(s))$ amortized time to delete the node in the heap and $O(1)$ time to delete the entry in the hash table.
 
-Otherwise, it takes O(d) amortized time to compute the new key for the hash table and O(log(s)) amortized time to update the heap given this key.
+Otherwise, it takes $O(d)$ amortized time to compute the new key for the hash table and $O(\log(s))$ amortized time to update the heap given this key.
 
-Thus, our algorithm requires O(log(s)) amortized time + expected O(1) time  + O(d) time to remove a hashtag from the Current-Window.
+Thus, our algorithm requires $O(\log(s))$ amortized time $+$ expected $O(1)$ time  $+ O(d)$ time to remove a hashtag from the Current-Window.
 
-The end-of-day updates to the History and the resulting updates to the heap take the following time.
-By Lemma 5 of the Hokusai paper, the amortized time required to do all end-of-day aggregation is O(md).
-Then, for each of the s nodes in the heap, it takes O(d) time to compute each updated key and O(log(s)) amortized time to update the heap given the new key.
+Now we analyze the time due to the end-of-day updates to the History and the resulting updates to the heap.
+By Lemma $5$ of the Hokusai paper, the amortized time required to do all end-of-day aggregation is $O(md)$.
+Then, for each of the $s$ nodes in the heap, it takes $O(d)$ time to compute each updated key and $O(\log(s))$ amortized time to update the heap given the new key.
 
-Thus, it takes O(md + slog(s)) amortized time + O(ds) time to do all necessary updates at the end of the day.
+Thus, it takes $O(md + slog(s))$ amortized time $+ O(ds)$ time to do all necessary updates at the end of the day.
 
-Querying the data structure for the top k trending items takes O(k*log(s)) amortized time for delete-max operations, O(k) time to announce that these items are trending, and O(k) amortized time to reinsert these nodes into the heap.
+Querying the data structure for the top $k$ trending items takes $O(k*log(s))$ amortized time for $\textbf{delete-max}$ operations, $O(k)$ time to announce that these items are trending, and $O(k)$ amortized time to reinsert these nodes into the heap.
 
-Thus, it takes O(k*log(s)) amortized time to determine what’s trending.
+Thus, it takes $O(k*log(s))$ amortized time to determine what is trending.
 
@@
 
-The History requires O(md) space for each Count-Min sketch, so it requires a total of O(md*log(T)) space.
+The History requires $O(md)$ space for each Count-Min sketch, so it requires a total of $O(md*\log(T))$ space.
+Each node in the heap requires a constant amount of space, so the heap requires $O(s)$ space.
+The hash table always contains at most $s$ entries with each entry requiring a constant amount of space.
+Also, in order to maintain an expected $O(1)$ lookup time, the hash table needs to have $O(s)$ bins.
+Thus, the hash table requires $O(s)$ space.
+The queue requires an entry for every hashtag still in the current window, so it requires $O(x)$ space.
 
-Each node in the heap requires a constant amount of space, so the heap requires O(s) space.
-
-The hash table always contains at most s entries with each entry requiring a constant amount of space.
-Also, in order to maintain an expected O(1) lookup time, the hash table needs to have O(s) bins.
-Thus, the hash table requires O(s) space.
-
-The queue requires an entry for every hashtag still in the current window, so it requires O(x) space.
-
-Thus, everything requires O(md*log(T) + x) space since s $<$ x.
+Thus, everything requires $O(md*\log(T) + x)$ space since $s < x$.
 
@@
 \subsection{Parameters of the History Data Structure}
-		pick some parameters here
-		n = 6, m = 3500 (size of hash tables), d = 20 (number of tables)
-		n: we do it for a month, so $\sim 2^{(6-1)}$ days is definitely enough
-		m = ceiling(e/epsilon), d = ceiling(ln(1/delta))
-		error in query is in factor epilson with probability 1 - delta
-		for d = 20, we have with high probability (1 - 1/$e^{20}$)
-		we have error factor of e/3500 of the total number of hashtags
-		total number of hashtags is $\sim$ 15 million
-		thus error bounds are in range $\sim$ 12000 on a given count
-		this error provides a 'good' amount of smoothing
+Since $T = 30$ (our data is for the month of September $2014$, therefore $\sim 2^{(6-1)}$ days is definitely enough), we store $6$ Count-Min
+sketches in the Hokusai portion of the History data structure, with an extra Count-Min 
+sketch for the Kernel.
+
+Each of these CM sketches has $m = 3500$ and $d = 20$. From the proof of CM sketch, we have that
+
+\begin{equation}
+\label{eq:error_control}
+m = \lceil \frac{e}{\epsilon} \rceil
+\end{equation}
+
+and
+
+\begin{equation}
+\label{eq:certainty_control}
+d = \lceil \textnormal{ln}(\frac{1}{\delta}) \rceil
+\end{equation}
+
+where the error in the query is in factor of $epsilon$ with probability $1 - \delta$.
+By Theorem $1$ in \cc{Matusevych:2012}, we have that if $n_x$ is the true count of a hashtag stored in the CM sketch, and $c_x$ is the estimate, we are guaranteed $n_x \leq c_x \leq n_x + \epsilon \sum_{x'} n_{x'}$.
+By choosing $m = 3500$, we have that $\epsilon = \frac{e}{3500}$ and $1 - \delta = 1 - \frac{1}{e^{20}}$, which is to say that we are within that error with very high probability.
+The error term is a multiplier on the total number of hashtags, of which there are $\sim 15$ million, as an upper bound.
+Thus we are within very high probability of the true count with range $\sim 12000$, on a given count.
+As a result, by storing less memory and keeping the memory down more, we get some amount of smoothing for free.
 
@@
 
-\subsection{Performance Measurements}
+\subsection{Runtime Measurements}
 
@@
 respective inputs to determine approximately how much space is used by each
-approach, and compare them.
-
-\subsubsection{Naive Algorithm}
-
-Space measurements (running on one week's worth of hashtags) (estimate 6 hours worth of hashtags)
-
-Time (running over one week): $\sim 15$ minutes
-
-\subsubsection{History-Sensitive Algorithm}
-
-(we had to build a history over the first three weeks of the month
- in order to determine the History data structure's values for the week we ran
- the algorithm over)
-
-Space measurements (running on one week's worth of hashtags):
-
-Time (running over one week): $2 - 3$ hours
+approach, and compare them. The naive algorithm's runtime on one week's worth
+of hashtags was $\sim 15$ minutes. Note that the naive algorithm did not have 
+to take into account the previous $3$ weeks, since it only ever depended on 
+the frequencies from $3$ hours before the present time. The history-sensitive
+algorithm on the other hand first had to build a history from the first three
+weeks of the month before running on the last week for comparison to the naive
+algorithm. The total running time for the history-sensitive algorithm was $
+\sim 2$ hours.
 
@@
 
-observed that history-sensitive 
-did about the same in terms of getting trending tweets (based on Jaccard), 
-but naive uses much less space.
+Based on the approximately normal, centered distribution of the Jaccard similarities, we observe that history-sensitive performed about as well as the naive approach in terms of getting trending tweets.
+However, the naive algorithm uses considerably less space and time (since we only need to check the
+$y$-unit before the current $y$-unit) to get similar results.
+Thus we conclude that the space efficiency of the naive approach outweighs the benefits of the history-sensitive approach for this particular task.
+is likely superior to the history-sensitive approach. Furthermore, in our development of the history-sensitive algorithm, utilized larger errors and smoothing to perform decently, suggesting that
+our approach to evaluating the history may not be the best one for tweet-trending detection.
+
+However, we note that from inspection the history-sensitive algorithm has more staying power than 
+the naive approach. While in the naive approach, it is rare for adjacent $y$-units to have the same
+trending tweets, it is a more common occurrence in the history-sensitive approach when it makes sense (for instance, there was an event going on for the whole day: the naive algorithm would immediately reduce showing the hashtags related to that event, while the history-sensitive algorithm would not discard that
+hashtag so easily).
 
@@
 
-do well to get month's worth of data in a few hours (2-3)
+In this paper, we only simulated our algorithm running on live data, since we did not 
+have good access to the live Twitter firehose. In terms of performance, despite being overshadowed
+by the simpler naive method due to their similar behavior (capturing a fair number of the same hashtags), 
+the history-sensitive algorithm did decently well. It took a few hours to run our algorithm 
+on a $15$ million hashtag dataset, with relatively large data structures to be updated and run in Python no less, a slower language.
 
-implementation wouldnt run in python, do it in C or something
+Were we to apply our algorithm to real-world real-time data, we would re-implement our code into C or another space and time efficient language.
 
-for real data from firehose (all of it), we would use 
-better infrastructure dedicated to processing all the data
-would use parallel setup (this scheme is adaptable to parallel
-setup)
+Furthermore, for real data from all of the Twitter firehose, we would require a better hardware infrastructure dedicated to processing all the data. Ideally we would use distributed systems and would
+parallelize our code before running it. The fact that the code is broken up into disparate parts make it easier to implement in parallel.
 
-seems reasonable if we have proper data storage setup.
+Therefore, given some adjustments, the algorithms developed in this paper open the doors to new exploration of topic modeling the Twitter data stream.
 
@@
 We provide a link to \href{https://github.com/kiranvodrahalli/cos521/}{all our code}, as well as to an online hosting of the \href{http://www.princeton.edu/~awlee/trends.html}{frequency visualization}.
+In the repository, there are also the results files, Powerpoint slides, and citations for the pictures we used.
 
@@
 Note that the heap priorities span a reasonable range from $0$ to $1$, and are not clustered around $1$ or $0$, implying the algorithms are actually differentiating the hashtags as trending and not trending.
+
+}
 
@@
 
-\bibitem[Image1]{Image1}
-\newblock BlowTheWhistle. Retrieved from \href{https://safesite.4agoodcause.com/curing-kids-cancer/donate/template_files/Blow_The_Whistle.jpg}{this link.}
-
-\bibitem[Image2]{Image2}
-\newblock How to Get Away with Murder.  Retrieved from \href{http://static.east.abc.go.com/service/image/index/id/65068ab1-d543-4958-a2cc-ee366bd807db/dim/600x800.jpg}{this link.}
-
-\bibitem[Image3]{Image3}
-\newblock One Direction terrorsist joke goes south for Daily Show.  Retrieved from \href{https://pbs.twimg.com/media/B52-mq-IQAEIqxz.jpg}{this link.}
-
-\bibitem[Image4]{Image4}
-\newblock Adrenna pic. Retrieved from \href{http://indiainf1.com/wp-content/uploads/2014/09/MRF-FF1600-BIC-28SEp2014-Adrenna-pic-300x184.jpg}{this link.}
-
-\bibitem[Image5]{Image5}
-\newblock Rory Mac.  Retrieved from \href{http://i.telegraph.co.uk/multimedia/archive/03050/rorymac_3050367b.jpg}{this link.}
-
-\bibitem[Image6]{Image6}
-\newblock Derek Jeter closes his Yankee Stadium finale with walk-off single.
-Retrieved from \href{http://www.foxsports.com/mlb/story/derek-jeter-closes-yankee-stadium-finale-with-walk-off-single-new-york-yankees-beat-baltimore-orioles-092514}{this link.}
-
 \end{thebibliography}

### 2015-01-13 paper.tex
minor paper changes

@@
 
-For each hashtag, the kernel sketch stores a positive value between $0$ and $2$ (though typically $\leq 1$) that approximates how often the hashtag showed up in the past.
+For each hashtag, the kernel sketch stores a positive value that approximates how often the hashtag showed up in the past.
 In \nref{sec:Correctness} we will show that this aggregate Count-Min sketch weights the hashtags that occurred $i > 0$ days ago with weight approximately $\frac{1}{i}$.
@@
 \end{proof}
-This value is approximately (frequency in last $y$-unit) / (freq current $z$-unit + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag $i$ $z$-units ago)).
+
+Given this proof and the fact that Count-Min sketches return approximate frequency counts, we know the priority function to be approximately:
+\begin{equation} 
+\label{eq:approxpriority}
+\textbf{priority}(\textnormal{hashtag}) \approx \frac{\textbf{freq}(\textnormal{hashtag}, \textnormal{present } y \textnormal{-unit})}{\textbf{freq}(\textnormal{hashtag}, \textnormal{present } z \textnormal{-unit}) + \sum\limits_{i=1}^T \frac{1}{i}* \textbf{freq}(\textnormal{hashtag}, i \ z \textnormal{-units ago})}
+\end{equation}
+%This value is approximately (frequency in last $y$-unit) / (freq current $z$-unit + $\sum\limits_{i=1}^T \frac{1}{i} *$ (frequency of hashtag $i$ $z$-units ago)).
 This seems to be a desirable function to maximize since it finds hashtags that are common in the last $y$-unit that have been comparatively infrequent in the past.
@@
 Thus, the hash table requires $O(s)$ space.
-The queue requires an entry for every hashtag still in the current window, so it requires $O(x)$ space.
+The queue requires an entry for every hashtag seen in the last $y$-unit, so it requires $O(x)$ space.
 

### 2015-01-13 paper.tex
Minor paper edits

@@
 We wanted to test whether an algorithm that uses
-the history (in condensed form) to help understand what is trending at the current time could outperform an algorithm with a very short term memory.
+the history (in condensed form) to better understand what is trending at the current time could outperform an algorithm with a very short term memory.
 
@@
 Now we analyze the time due to the end-of-day updates to the History and the resulting updates to the heap.
-By Lemma $5$ of the Hokusai paper, the amortized time required to do all end-of-day aggregation is $O(md)$.
+By Lemma $5$ in \cc{Matusevych:2012}, the amortized time required to do all end-of-day aggregation is $O(md)$.
 Then, for each of the $s$ nodes in the heap, it takes $O(d)$ time to compute each updated key and $O(\log(s))$ amortized time to update the heap given the new key.
 
-Thus, it takes $O(md + slog(s))$ amortized time $+ O(ds)$ time to do all necessary updates at the end of the day.
+Thus, it takes $O(md + s\log(s))$ amortized time $+ O(ds)$ time to do all necessary updates at the end of the day.
 
-Querying the data structure for the top $k$ trending items takes $O(k*log(s))$ amortized time for $\textbf{delete-max}$ operations, $O(k)$ time to announce that these items are trending, and $O(k)$ amortized time to reinsert these nodes into the heap.
+Querying the data structure for the top $k$ trending items takes $O(k\log(s))$ amortized time for $\textbf{delete-max}$ operations, $O(k)$ time to announce that these items are trending, and $O(k)$ amortized time to reinsert these nodes into the heap.
 
-Thus, it takes $O(k*log(s))$ amortized time to determine what is trending.
+Thus, it takes $O(k\log(s))$ amortized time to determine what is trending.
 
@@
 
-The History requires $O(md)$ space for each Count-Min sketch, so it requires a total of $O(md*\log(T))$ space.
+The History requires $O(md)$ space for each Count-Min sketch, so it requires a total of $O(md\log(T))$ space.
 Each node in the heap requires a constant amount of space, so the heap requires $O(s)$ space.
@@
 
-Thus, everything requires $O(md*\log(T) + x)$ space since $s < x$.
+Thus, everything requires $O(md\log(T) + x)$ space since $s < x$.
 
@@
 For $z$, the unit time period for the History data structure, we chose day-length intervals.
-These values made sense because the testing data we used occurred over a $30$-day interval, and we wanted to avoid too fine a temporal resolution so that we could run our code on the data in reasonable time.
+These values made sense because the testing data we used occurred over a $30$-day interval, and we wanted to avoid too fine a temporal resolution so that we could run our code on the data in a reasonable amount of time.
 
@@
 
-where the error in the query is in factor of $epsilon$ with probability $1 - \delta$.
+where the error in the query is in factor of $\epsilon$ with probability $1 - \delta$.
 By Theorem $1$ in \cc{Matusevych:2012}, we have that if $n_x$ is the true count of a hashtag stored in the CM sketch, and $c_x$ is the estimate, we are guaranteed $n_x \leq c_x \leq n_x + \epsilon \sum_{x'} n_{x'}$.
@@
 Thus we are within very high probability of the true count with range $\sim 12000$, on a given count.
-As a result, by storing less memory and keeping the memory down more, we get some amount of smoothing for free.
+
+By using less memory we increase the amount of error.
+These error terms are added to the denominator of our priority function, and as a result, they prevent hashtags with very relatively low frequency from being identified as trending, which is a very desirable property.
 
@@
 We downloaded the \href{https://archive.org/details/twitterstream}{September $2014$ archive of Twitter data } to serve as test data for our algorithms \cc{Twitter2014}.
-We then parsed the data into (timestamp, hashtag) format
+We then parsed the data into (timestamp, hashtag) pairs
 for all thirty days worth of data.
@@
 
-We plot it below in the histogram shown in \aref{fig:jaccard_sims}.
+We plot this data in the histogram shown in \aref{fig:jaccard_sims}.
 The $x$-axis is the Jaccard similarity, and the $y$-axis is the count.
@@
 at around the time our algorithms say the hashtag relating to the news
-begins to trend. A fair number of the trending hashtags related to sports and TV shows.
+begins to trend. A fair number of the trending hashtags are related to sports and TV shows.
 
@@
 
-Based on the approximately normal, centered distribution of the Jaccard similarities, we observe that history-sensitive performed about as well as the naive approach in terms of getting trending tweets.
+Based on the approximately normal, centered distribution of the Jaccard similarities, we observe that the history-sensitive approach performed about as well as the naive approach in terms of getting trending tweets.
 However, the naive algorithm uses considerably less space and time (since we only need to check the
@@
 Thus we conclude that the space efficiency of the naive approach outweighs the benefits of the history-sensitive approach for this particular task.
-is likely superior to the history-sensitive approach. Furthermore, in our development of the history-sensitive algorithm, utilized larger errors and smoothing to perform decently, suggesting that
-our approach to evaluating the history may not be the best one for tweet-trending detection.
+Furthermore, in our development of the history-sensitive algorithm, we utilized larger errors as a method of smoothing to perform decently, suggesting that our approach for evaluating the history may not be the best one for tweet-trending detection.
 
@@
 Furthermore, for real data from all of the Twitter firehose, we would require a better hardware infrastructure dedicated to processing all the data. Ideally we would use distributed systems and would
-parallelize our code before running it. The fact that the code is broken up into disparate parts make it easier to implement in parallel.
+parallelize our code before running it. The fact that the code is broken up into disparate parts makes it easier to implement in parallel.
 
@@
 
-We also know that an estimation the top $k$ hashtags in a given time interval could provide a topic model for tweets.
+We also know that an estimation of the top $k$ hashtags in a given time interval could provide a topic model for tweets.
 It could be interesting to use the top $k$ hashtags to build a time-sensitive topic model and analyze quantities including the speed of topic change, the variety of topics, and so on.
 
-Even though not perhaps irrelevant to the specific task of identifying trending hashtags, 
+Even though it is perhaps not suited for the specific task of identifying trending hashtags, 
 the approach of the history-sensitive algorithm may be useful in other domains.

### 2015-01-13 viz/trending.js
Add time for viz

@@ -22,6 +22,22 @@ var trendingChart = new Chart(ctx).Bar(data, {
     scaleFontColor: '#eee'
 });
 
+
+var STEP_INCR = 3; // 3 hrs
+var curDate = new Date(2014, 8, 24);
+var dateToString = function(d) {
+    var fields = d.toString().split(':');
+    return fields[0] + ':' + fields[1];
+};
+var prevSliderVal = 0;
+
+Date.prototype.addHours = function(h) {
+    this.setHours(this.getHours() + h);
+    return this;
+};
+
+$('#time').text(dateToString(curDate));
+
 $( "#slider" ).slider({
     value: 0,
     min: 0,
@@ -34,5 +50,8 @@ $( "#slider" ).slider({
 	    trendingChart.datasets[0].bars[i].value = allData[curDataIdx][i];
 	}
 	trendingChart.update();
+	curDate.addHours((ui.value - prevSliderVal) * STEP_INCR);
+	$('#time').text(dateToString(curDate));
+	prevSliderVal = ui.value;
     }
 });

### 2015-01-13 paper.tex
Final paper changes

@@
 
+Because Twitter limits access to its data, the datasets on which we tested our algorithms only contained 1\% of the total number of tweets.
+If we were to run our algorithms on the entire dataset, we would expect that the space efficiency of the history-sensitive algorithm would improve relative to the naive algorithm because the naive algorithm would need to store 100 times more data while the history-sensitive algorithm could just reduce the size of the $y$-units to keep it nearly as space efficient.
+We would not expect this to greatly affect the history-sensitive algorithm's accuracy because we could still process the same number of recent tweets to compare against the History structure.
+Thus, on the full dataset, we could still reasonably expect the history-sensitive algorithm to be at least as space efficient as the naive algorithm.
+
 \subsection{Applicability to Real-Time Data}

### 2015-01-13 paper.tex
minor fixes
updated paper more

@@
 Since its founding in $2006$, Twitter has grown into a corporation dominating a significant
-proporition of social media today. In recent years, it has become interesting to analyze tweets
+proportion of social media today. In recent years, it has become interesting to analyze tweets
 to determine interesting phenomena ranging from the viral spread of news to the detection of earthquakes \cc{Burks:2014}.
@@
 
-More specifically, we will attempt to approximate the $k$ most popular Twitter hashtags in the time intervals of length $y$ in order to tell what is trending during that time period.
+More specifically, we will attempt to approximate the $k$ most popular Twitter hashtags in time intervals of length $y$ in order to tell what is trending during that time period.
 The idea is that estimating the top $k$ hashtags in some time interval provides a model for the topics that are trending on Twitter in that time period.
@@
 Twitter itself stores all of its frequency data, and looks at recent spikes while conditioning on all of the past frequency data in order to determine estimates of trending likelihood.
-In $2012$, MIT Professor Devavrat Shah built a machine-learning algorithm to predict which topics will trend on Twitter an hour and a half before they do, with $95\%$ accuracy, though the datasets they tested their algorithm on were small\cc{MITNews2012}.
+In $2012$, MIT Professor Devavrat Shah built a machine-learning algorithm to predict which topics will trend on Twitter an hour and a half before they do, with $95\%$ accuracy, though the datasets they tested their algorithm on were small \cc{MITNews2012}.
 
@@
 The reason why the history is important is as follows: Some hashtags are constantly present.
-For example, in our experiments, the hashtags $\#$porn, $\#$gameinsight, $\#$teamfollowback had nearly always had a high frequency in the current time unit.
+For example, in our experiments, the hashtags $\#porn$, $\#gameinsight$, $\#teamfollowback$ had nearly always had a high frequency in the current time unit.
 This fact is not surprising given the nature of the first hashtag.
@@
 The rule is if you follow that tag, you follow people who post that tag.
-Since these sorts of tags are consistently present, they should not be consider trending: they are not new.
+Since these sorts of tags are consistently present, they should not be considered trending: they are not new.
 Therefore, we need to filter these out with the history.
@@
 To store data about all occurrences of every hashtag seen in the past, we use a modified version of the time-aggregated Hokusai system \cc{Matusevych:2012}, which is an extension of the Count-Min sketch.
-We previously described this data structure in \nref{sec:PreviousWork}.
+We already described this data structure in \nref{sec:PreviousWork}.
 To the Hokusai structure we add another Count-Min sketch that combines the information from the Hokusai Count-Min sketches.
@@
 Its role is to depreciate the value of older data.
-Denoting the CM sketches of the Hokusai structure as a vector $\textbf{M} = \{\overline{M}, M^0, M^1, ..., M^{{\lfloor {\log(T)} \rfloor}}\}$, where $\overline{M}$ is the Count-Min sketch storing the data of a unit time step (of size $z$), $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ time resolution, and $T$ is an upper bound on the number of $z$-units stored by the data structure.
+We denote the CM sketches of the Hokusai structure as a vector $\textbf{M} = \{\overline{M}, M^0, M^1, ..., M^{{\lfloor {\log(T)} \rfloor}}\}$, where $\overline{M}$ is the Count-Min sketch storing the data of a unit time step (of size $z$), $M^j$ is the Count-Min sketch storing the data for the sketch with a $2^j$ time resolution, and $T$ is an upper bound on the number of $z$-units stored by the data structure.
 Then, the kernel function is given by
@@
 \While {data arrives}
-	\State Aggregate data into sketch $\overline{M}$ for current $z$-unit while also adding this data to $K$
+	\State Aggregate data into sketch $\overline{M}$ for current $z$-unit
 	\State $t \leftarrow t + 1$ (increment counter)
-	\State $K \leftarrow K - \overline{M}$
 	\For {$j = 0$ to argmax \{$l$ where $t$ mod $2^l = 0$\}}
-		\State $K \leftarrow K + 2^{-j}(\overline{M} - M^j)$
+		\State $K \leftarrow K + 2^{-j}(\overline{M} - M^j)$ (update the Kernel)
 		\State $T \leftarrow \overline{M}$ (back up temporary storage)
@@
 			\State look up the hashtag corresponding to this node in the hash table
-			\State Update the key of the node to: \[\frac{\textnormal{frequency in last } y-\textnormal{length time period found at the table entry}}{\textnormal{new value in History data structure}}\]
+			\State Update the key of the node to: $\textbf{priority}(\textnormal{hashtag})$
 		\EndFor
@@
 			\Else	
-				\State Update the key of this node pointed to by this entry in the table to the proper value given the new frequency.
+				\State Update the key of the node with the new frequency.
 			\EndIf
@@
 The $x$-axis is the Jaccard similarity, and the $y$-axis is the count.
-The mean was $0.47$, and the standard Deviation was $0.18$.
+The mean was $0.47$, and the standard deviation was $0.18$.
 From the plot, we can see that the distribution of the Jaccard similarities is roughly a centered normal distribution around $0.5$.
@@
 
-Therefore, given some adjustments, the algorithms developed in this paper open the doors to new exploration of topic modeling the Twitter data stream.
+Therefore, given some adjustments, the algorithms developed in this paper open the doors to new explorations of topic modeling the Twitter data stream.
 
@@
 We would also like to try different kernel functions to see how they performed.
-Another approach to utilizing the history information would be instead of weighting by inverse exponentially weighted frequency in the past for a given hashtag, attempt to calculate some score of periodicity for the hashtag.
-If the hashtag has low periodicity over a given time scale, then even if it appears multiple times in the past (because the past is long), our algorithm will avoid discarding it as non-trending.
+Another approach to utilizing the history information would be instead of weighting by inverse-exponentially-weighted frequency in the past for a given hashtag, attempt to calculate some score of periodicity for the hashtag.
+If the hashtag has low periodicity over a given time scale, then even if it appears multiple times in the past, our algorithm will avoid discarding it as non-trending despite the potentially high frequency count (because the past is long).
 
@@
 
-We provide a link to \href{https://github.com/kiranvodrahalli/cos521/}{all our code}, as well as to an online hosting of the \href{http://www.princeton.edu/~awlee/trends.html}{frequency visualization}.
-In the repository, there are also the results files, Powerpoint slides, and citations for the pictures we used.
+We provide a link to \href{https://github.com/kiranvodrahalli/cos521/}{all our code}. In the repository, there are also the results files, Powerpoint slides, and citations for the pictures we used.
+
+We also provide a link to an online hosting of the \href{http://www.princeton.edu/~awlee/trends.html}{frequency visualization}.
+The frequency visualization is a histogram of hashtag frequencies, displayed over time.
+We cherry-picked a set of interesting hashtags based on our observations, and then made this visualization to present the changes in frequency in $3$-hour intervals over time of these hashtags.
+
+The fact that these frequencies change over time and spike and fall in turns suggests that
+when the hashtag frequencies spike, the hashtags are trending. The goal of our algorithms is to identify the hashtags (of a much larger set) which are trending at a given point in time.
+
+So we hope that for a selected $3$-hour interval, the hashtags that our algorithms say are trending include the ones that are trending at that time in this visualization.
 
@@
 
-\section{Appendix II: Top-$k$ Hashtags in $6$ Interesting Time Chunks} \label{sec:Appendix_topk}
+\section{Appendix II: Top-$k$ Hashtags Results} \label{sec:Appendix_topk}
 
